---
title: "Add Ecozone"
author: "Anna Talucci"
date: "2023-07-11"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```


# Overview
Create ecoregion centroids for Manuscript map.



Notes:
[Better colors for Mapping](https://www.esri.com/arcgis-blog/products/js-api-arcgis/mapping/better-colors-for-better-mapping/)
# Packages
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
```

# Projection
[Some projection info]https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection)

```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```

# Data

## Points

```{r}
points = st_read("../data/allPoints/AllDataPts.shp", "AllDataPts")
```

## WWF
```{r}
wwf = st_read("../data/wwfEcozones/wwf_terr_ecos.shp", "wwf_terr_ecos")
```


# WWF Data 
```{r}
wwf 
```
```{r}
wwf %>% st_drop_geometry() %>% distinct(REALM)
```


# Select Realms
```{r}
arctic = wwf %>% filter(REALM %in% c("PA", "NA")) %>% dplyr::select(OBJECTID, ECO_NAME, REALM, BIOME, ECO_NUM, ECO_ID, geometry)
```

```{r}
arctic
```
```{r eval=FALSE, include=FALSE}
plot(st_geometry(arctic))
```



# Add Ecoregions to points

```{r}
combineData = st_join(points, arctic) 
```

```{r}
combineData
```
```{r}
unique(combineData$ECO_NAME)
```
```{r}
cat(paste0(sprintf('"%s"', sort(unique(combineData$ECO_NAME))), collapse = ", "))
```


# Create Centroid by Ecoregion
```{r}
( ecoPoly = arctic %>% 
    filter(ECO_NAME %in% c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "Beringia upland tundra", "Brooks-British Range tundra", "Cherskii-Kolyma mountain tundra", "Chukchi Peninsula tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Pacific Coastal Mountain icefields and tundra", "Yamal-Gydan tundra")) )
```

## 1. "Arctic coastal tundra"
```{r}
eco01 =  arctic %>% 
    filter(ECO_NAME=='Arctic coastal tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Arctic coastal tundra')
```

## 2. "Arctic foothills tundra"
```{r}
eco02 =  arctic %>% 
    filter(ECO_NAME=='Arctic foothills tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Arctic foothills tundra')
```

## 3. "Beringia lowland tundra"
```{r}
eco03 =  arctic %>% 
    filter(ECO_NAME=='Beringia lowland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Beringia lowland tundra')
```
## 4. "Beringia upland tundra" 

```{r}
eco04 =  arctic %>% 
    filter(ECO_NAME=='Beringia upland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Beringia upland tundra')
```

## 5. "Brooks-British Range tundra"

```{r}
eco05 =  arctic %>% 
    filter(ECO_NAME=='Brooks-British Range tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Brooks-British Range tundra')
```

## 6. "Cherskii-Kolyma mountain tundra"

```{r}
eco06 =  arctic %>% 
    filter(ECO_NAME=='Cherskii-Kolyma mountain tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Cherskii-Kolyma mountain tundra')
```
## 7. "Chukchi Peninsula tundra"

```{r}
eco07 =  arctic %>% 
    filter(ECO_NAME=='Chukchi Peninsula tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Chukchi Peninsula tundra')
```
## 8. "East Siberian taiga"
```{r}
eco08 =  arctic %>% 
    filter(ECO_NAME=='East Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='East Siberian taiga')
```

## 9. "Interior Alaska-Yukon lowland taiga"
```{r}
eco09 =  arctic %>% 
    filter(ECO_NAME=='Interior Alaska-Yukon lowland taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Interior Alaska-Yukon lowland taiga')
```

## 10. "Interior Yukon-Alaska alpine tundra"
```{r}
eco10 =  arctic %>% 
    filter(ECO_NAME=='Interior Yukon-Alaska alpine tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Interior Yukon-Alaska alpine tundra')
```

## 11. "Muskwa-Slave Lake forests"
```{r}
eco11 =  arctic %>% 
    filter(ECO_NAME=='Muskwa-Slave Lake forests') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Muskwa-Slave Lake forests')
```

## 12. "Northeast Siberian taiga"
```{r}
eco12=  arctic %>% 
    filter(ECO_NAME=='Northeast Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northeast Siberian taiga')
```

## 13. "Northern Canadian Shield taiga"
```{r}
eco13=  arctic %>% 
    filter(ECO_NAME=='Northern Canadian Shield taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northern Canadian Shield taiga')
```
## 14. "Northwest Territories taiga"
```{r}
( eco14=  arctic %>% 
    filter(ECO_NAME=='Northwest Territories taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northwest Territories taiga')
)
```

## 15. "Pacific Coastal Mountain icefields and tundra" 

```{r}
( eco15 = arctic %>% 
    filter(ECO_NAME=='Pacific Coastal Mountain icefields and tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Pacific Coastal Mountain icefields and tundra')
)
```
## 16. "Yamal-Gydan tundra"

```{r}
( eco16 =  arctic %>% 
    filter(ECO_NAME=='Yamal-Gydan tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Yamal-Gydan tundra')
)
```
```{r}
( 
  ecocentroid = bind_rows(eco01, eco02, eco03, eco04, eco05, eco06, eco07, eco08, eco09, eco10, eco11, eco12, eco13, eco14, eco15, eco16) %>% rename(geometry = x)
)
```


```{r}
( reproj = ecocentroid %>% st_transform(., crs = polarProj) %>% 
    dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]))


```



# grouped points
```{r}
(eco_groupedCt <- combineData %>% st_drop_geometry() %>%
                group_by(ECO_NAME) %>%
                summarise(count = n()) )
```
```{r}
( eco_grouped <- combineData %>%
                group_by(ECO_NAME) %>%
                summarise(st_union(geometry)) %>%
                st_centroid() %>%
  left_join(., eco_groupedCt, by= 'ECO_NAME') %>%
  mutate(total = sum(count)) %>%
  mutate(prop = round(count/total, 5)) %>%
  mutate(percent = round(prop*100, 2)) %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() )

```


```{r}
(
  EcoXY = reproj %>%
  left_join(., eco_grouped, by='ECO_NAME') %>%
  mutate(prct = ifelse(percent ==0.38, 0.4,
                ifelse(percent ==42.53, 42.5,
                ifelse(percent==4.49, 4.5,
                ifelse(percent==0.08, 0.08,
                ifelse(percent==0.06, 0.06,
                ifelse(percent==0.05, 0.05,
                ifelse(percent==0.95, 0.95,
                ifelse(percent==0.34, 0.3,
                ifelse(percent==12.81, 12.8,
                ifelse(percent==6.10, 6.1,
                ifelse(percent==22.44, 22.4,
                ifelse(percent==1.5, 1.5,
                ifelse(percent==2.16, 2.2,
                ifelse(percent==3.08, 3.1,
                       ifelse(percent==2.76, 2.8,
                              ifelse(percent==0.01, 0.01,
                                     ifelse(percent==0.26, 0.3, -99))))))))))))))))))
)

```

```{r}
(
  EcoXY = reproj %>%
  left_join(., eco_grouped, by='ECO_NAME') %>%
  mutate(prct = signif(percent,digits=3))
)

```
```{r}
EcoXY
```
### write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(combineData, "../outputs/points/PtsEcozones.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(eco_grouped , "../outputs/points/PtsEcoSum.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(EcoXY , "../outputs/points/PtsEcoSumXYResolve.shp", driver="ESRI Shapefile")
```