---
title: "Add Ecozone"
author: "Anna Talucci"
date: "2023-07-11"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```


# Overview

Notes:
[Better colors for Mapping](https://www.esri.com/arcgis-blog/products/js-api-arcgis/mapping/better-colors-for-better-mapping/)
# Packages
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
```
# Projection
```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```

# Data

## Points

```{r}
points = st_read("../data/allPoints/AllDataPts.shp", "AllDataPts")
```

## WWF
```{r}
wwf = st_read("../data/wwfEcozones/wwf_terr_ecos.shp", "wwf_terr_ecos")
```


# WWF Data 
```{r}
wwf 
```
```{r}
wwf %>% st_drop_geometry() %>% distinct(REALM)
```


# Select Realms
```{r}
arctic = wwf %>% filter(REALM %in% c("PA", "NA")) %>% dplyr::select(OBJECTID, ECO_NAME, REALM, BIOME, ECO_NUM, ECO_ID, geometry)
```

```{r}
arctic
```
```{r}
plot(st_geometry(arctic))
```



# Add Ecoregions to points

```{r}
combineData = st_join(points, arctic) 
```

```{r}
combineData
```
# Eco region centroids
'Arctic coastal tundra', 'Arctic foothills tundra', 'Beringia lowland tundra', 'East Siberian taiga', 'Interior Alaska-Yukon lowland taiga', 'Interior Yukon-Alaska alpine tundra', 'Muskwa-Slave Lake forests', 'Northeast Siberian taiga', 'Northern Canadian Shield taiga', 'Northwest Territories taiga'
# Projection
[Some projection info]https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection)
```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```
```{r}
( ecoPoly = arctic %>% 
    filter(ECO_NAME %in% c('Arctic coastal tundra', 'Arctic foothills tundra', 'Beringia lowland tundra', 'East Siberian taiga', 'Interior Alaska-Yukon lowland taiga', 'Interior Yukon-Alaska alpine tundra', 'Muskwa-Slave Lake forests', 'Northeast Siberian taiga', 'Northern Canadian Shield taiga', 'Northwest Territories taiga')) )
```

st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() # make it a sf object again
```{r}
ArcCoastTund =  arctic %>% 
    filter(ECO_NAME=='Arctic coastal tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Arctic coastal tundra')
```

```{r}
ArcFootTund =  arctic %>% 
    filter(ECO_NAME=='Arctic foothills tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Arctic foothills tundra')
```

```{r}
BerLowTund =  arctic %>% 
    filter(ECO_NAME=='Beringia lowland tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Beringia lowland tundra')
```

```{r}
EastSibTaig =  arctic %>% 
    filter(ECO_NAME=='East Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='East Siberian taiga')
```

```{r}
InAkYkLowTai =  arctic %>% 
    filter(ECO_NAME=='Interior Alaska-Yukon lowland taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Interior Alaska-Yukon lowland taiga')
```

```{r}
InYkAkAlpTai =  arctic %>% 
    filter(ECO_NAME=='Interior Yukon-Alaska alpine tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Interior Yukon-Alaska alpine tundra')
```

```{r}
MuckSlaLkF =  arctic %>% 
    filter(ECO_NAME=='Muskwa-Slave Lake forests') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Muskwa-Slave Lake forests')
```

```{r}
NESTai=  arctic %>% 
    filter(ECO_NAME=='Northeast Siberian taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northeast Siberian taiga')
```

```{r}
CanTaigSh=  arctic %>% 
    filter(ECO_NAME=='Northern Canadian Shield taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northern Canadian Shield taiga')
```

```{r}
( NWTTaig=  arctic %>% 
    filter(ECO_NAME=='Northwest Territories taiga') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Northwest Territories taiga')
)
```
```{r}
( ygTun=  arctic %>% 
    filter(ECO_NAME=='Yamal-Gydan tundra') %>%
    st_geometry() %>% # pull just geometry
  st_combine() %>%  # from many points to a single multipoint
  st_centroid() %>% # compute centroid of the multipoint
  st_as_sf() %>% # make it a sf object again 
  mutate(ECO_NAME='Yamal-Gydan tundra')
)
```
```{r}
( 
  ecocentroid = bind_rows(ArcCoastTund, ArcFootTund, BerLowTund, EastSibTaig, InAkYkLowTai, InYkAkAlpTai, MuckSlaLkF, NESTai, CanTaigSh, NWTTaig, ygTun) %>% rename(geometry = x)
)
```


```{r}
( reproj = ecocentroid %>% st_transform(., crs = polarProj) %>% 
    dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]))


```



# grouped points
```{r}
(eco_groupedCt <- combineData %>% st_drop_geometry() %>%
                group_by(ECO_NAME) %>%
                summarise(count = n()) )
```
```{r}
eco_grouped <- combineData %>%
                group_by(ECO_NAME) %>%
                summarise(st_union(geometry)) %>%
                st_centroid() %>%
  left_join(., eco_groupedCt, by= 'ECO_NAME') %>%
  mutate(total = sum(count)) %>%
  mutate(prop = round(count/total, 5)) %>%
  mutate(percent = round(prop*100, 2)) %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()
```


```{r}
EcoXY = reproj %>%
  left_join(., eco_grouped, by='ECO_NAME') %>%
  mutate(prct = ifelse(percent ==0.41, 0.4,
                ifelse(percent ==45.32, 45.3,
                ifelse(percent==6.61, 6.6,
                ifelse(percent==0.73, 0.7,
                ifelse(percent==18.41, 18.4,
                ifelse(percent==1.65, 1.7,
                ifelse(percent==23.34, 23.3,
                ifelse(percent==1.43, 1.4,
                       ifelse(percent==0.58, 0.6,
                              ifelse(percent==1.52, 1.5,
                                     ifelse(percent==0.01, 0.01, -99))))))))))))

```

```{r}
EcoXY
```
### write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(combineData, "../outputs/points/PtsEcozones.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(eco_grouped , "../outputs/points/PtsEcoSum.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(EcoXY , "../outputs/points/PtsEcoSumXY.shp", driver="ESRI Shapefile")
```