---
title: " Gibson/Olefeldt Clean to Point"
author: "Anna Talucci"
date: '2022-12-11'
output: html_document
---

Gibson/Olefeldt


# clear environment
```{r}
rm(list=ls())
```

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Olefeldt
```{r}
f12 <- read.csv(f[12], header = TRUE)
f12
```

# fix incorrect logical columns
```{r}
# fix incorrect logical columns
f12$slope <- read.csv(f[12], header = TRUE, colClasses = "character")[,20]

# remove non-numeric characters from numeric vars
# note we're loosing info on where Organic Layer Thickness is in excess of the entered value
f12$organic_depth <- as.numeric(gsub(">", "",f12$organic_depth))

f12 %>% distinct(site_id,year,month,day,fire_id,burn_unburn)

```

```{r}
str(f12)
```

```{r}
f12_dates = f12 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) #%>%
  #dplyr::select(-fire_id)

f12_dates
```


```{r}
f12_dates %>% filter_all(any_vars(is.na(.))) 
```

```{r}
f12_dates %>% distinct(site_id, plot_id, burn_unburn) %>% write.csv(., '../outputs/pairs/olefeldt.csv')
```

```{r}
f12_dates %>% distinct(julianDate, year, date)
```


# Finish cleaning to match other datasets
```{r}
( 
  f12_clean = f12_dates %>%  
  mutate(across(c(plot_id, site_id, fire_id), as.character)) %>% 
  mutate(across(c(fire_year, organic_depth, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) %>%
  rename(lastNm = last_name, cntryId = country_code, lon=long, biome=boreal_tundra, vegCC = veg_cover_class, siteId = site_id, plotId = plot_id, fireId = fire_id, fireYr = fire_year, distur = burn_unburn, orgDpth = organic_depth, msrDepth = thaw_depth, hitRock=hit_rock, gtProbe = gt_probe, msrType = thaw_active, topoPos = topo_position, jDate = julianDate) %>%
    mutate(msrType = as.character(msrType))
)
```

# Create points shapefile

## For f12
sf::st_as_sf(dd, coords = c("x","y"))
```{r}
f12_pts = st_as_sf(f12_clean, coords = c("lon","lat"), crs = 4326, remove = FALSE)
```


### write to shapefile
```{r}
st_write(f12_pts, "../outputs/allPoints/Olefeldt_pts.shp", driver="ESRI Shapefile")
```