---
title: "Veraverbeke Clean data to point"
author: "Anna Talucci"
date: '2022-12-02'
output: html_document
---


# clear environment
```{r}
rm(list=ls())
```

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Veraverbeke
```{r}
f16 <- read.csv(f[16], header = TRUE)
f16
```
```{r}
str(f16)
```
# fix incorrect logical columns
```{r}
f16$thaw_active <- read.csv(f[16], header = TRUE, colClasses = "character")[,19]
```

```{r}
f16_dates = f16 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) 

f16_dates
```

```{r}
f16_dates %>% distinct(site_id, plot_id, fire_id, burn_unburn) %>% write.csv(., '../outputs/pairs/veraverbeke.csv')
```
```{r}
f16_dates %>% distinct(julianDate, year, date)
```

# Split by active & thaw
```{r}
(
  f16_all = f16_dates %>% 
  drop_na(long) %>%
  drop_na(lat)
)

( 
  f16_thaw = f16_dates %>% 
  filter(thaw_active == "T") 
)



(
  f16_active = f16_dates %>% 
  filter(thaw_active == "A") 
)
```

# Create points shapefile

## For f16
sf::st_as_sf(dd, coords = c("x","y"))
```{r}
f16_thaw_pts = st_as_sf(f16_thaw, coords = c("long","lat"), crs = 4326, remove = FALSE)
f16_pts = st_as_sf(f16_all, coords = c("long","lat"), crs = 4326, remove = FALSE)
```


### write to shapefile
```{r}
st_write(f16_thaw_pts, "../outputs/ThawPoints/Veraverbeke_thaw_pts.shp", driver="ESRI Shapefile")

st_write(f16_pts, "../outputs/allPoints/Veraverbeke_pts.shp", driver="ESRI Shapefile")
```