---
title: "Buma Clean to point"
author: "Anna Talucci"
date: '2022-12-07'
output: html_document
---

**MISING DAY OF MONTH for all thaw depth measures (n=340 which is all observations)**

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Buma
# read without 17th column, which is redundant to the gt_prob column
```{r}
f3 <- read.csv(f[3], header = TRUE)[,-17]
```
# fix thaw/active column, which was read as logical
```{r}
f3$thaw_active <- read.csv(f[3], header = TRUE, colClasses = "character")[,20]
```
# fix site_id for aggregating (see Note below from Buma)
```{r}
f3$site_id <- paste(f3$site_id, sapply(strsplit(f3$plot,"_"),"[[",2), sep="_" )
```

# fix positive longitude values, which are in the wrong hemisphere
```{r}
f3$long <- ifelse(f3$long > 0, -f3$long,f3$long)
```
NOTE---The Dalton and the Steese are larger "sites," each of which has a lot of plots in them.  But those plots do vary, so don't aggregate.  At the Dalton, there are sites with 0 (unburned), 1 (one fire, which would be 2004/2005 era), 2 fires (which would be 1970's era AND 2004 or 2005), or 3 fires (which would be 1950's, 1970's, and 2004 or 2005.  So if aggregating, what you'd want to do would be to aggregate by those treatments (0, 1, 2, or 3 fires) within the Dalton or Steese "sites."  So, sounds like you'd want to aggregate all the unburned plots at the Dalton, all the 1 burn plots at the Dalton, etc.  I would not aggregate Dalton and Steese plots together, they are functionally different sites (uplands vs. low lands, respectively).

```{r}
f3
```

```{r}
f3 %>% distinct(site_id,year,month,day,fire_id,burn_unburn)
```

```{r}
str(f3)
```
```{r}
f3 %>% 
  group_by(thaw_active) %>%
  count(.)
```

```{r}
f3_dates = f3 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) 

f3_dates
```

```{r}
f3_dates %>% distinct(site_id, burn_unburn) 
```

```{r}
f3_dates %>% distinct(site_id, plot_id, fire_id, burn_unburn) %>% write.csv(., '../outputs/pairs/buma.csv')
```

```{r}
f3_dates %>% distinct(julianDate, year, date)
```

# Finish cleaning to match other datasets
```{r}
( 
  f3_clean = f3_dates %>%  
  mutate(across(c(plot_id, site_id, fire_id), as.character)) %>% 
  mutate(across(c(fire_year, organic_depth, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) %>%
  rename(lastNm = last_name, cntryId = country_code, lon=long, biome=boreal_tundra, vegCC = veg_cover_class, siteId = site_id, plotId = plot_id, fireId = fire_id, fireYr = fire_year, distur = burn_unburn, orgDpth = organic_depth, msrDepth = thaw_depth, hitRock=hit_rock, gtProbe = gt_probe, msrType = thaw_active, topoPos = topo_position, jDate = julianDate) %>%
    mutate(msrType = as.character(msrType))
)
```


# Create points shapefile

## For F3
```{r}
f3_pts = st_as_sf(f3_clean, coords = c("lon","lat"), crs = 4326, remove = FALSE)
```

### write to shapefile
```{r}
st_write(f3_pts, "../outputs/allPoints/Buma_pts.shp", driver="ESRI Shapefile")
```
