---
title: "Combine all point files for mapping"
author: "Anna Talucci"
date: "2023-03-22"
output: html_document
---
# clear environment
```{r}
rm(list=ls())
```

# Overview

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data
```{r}
f01 = st_read("../data/allPoints/Baillargeon_pts.shp", "Baillargeon_pts") 
f02 = st_read("../data/allPoints/Buma_pts.shp", "Buma_pts") 
f03 = st_read("../data/allPoints/Dielman_pts.shp", "Dielman_pts") 
f04 = st_read("../data/allPoints/Douglas_pts.shp", "Douglas_pts") 
f05 = st_read("../data/allPoints/Frost_pts.shp", "Frost_pts") 
f06 = st_read("../data/allPoints/Gaglioti_pts.shp", "Gaglioti_pts") 
f07 = st_read("../data/allPoints/Manies_pts.shp", "Manies_pts") 
f08 = st_read("../data/allPoints/Natali_pts.shp", "Natali_pts") 
f09 = st_read("../data/allPoints/ODonnell_pts.shp", "ODonnell_pts") 
f10 = st_read("../data/allPoints/Olefeldt_pts.shp", "Olefeldt_pts") 
f11 = st_read("../data/allPoints/Paulson_pts.shp", "Paulson_pts")
f12 = st_read("../data/allPoints/Rocha_pts.shp", "Rocha_pts") 
f13 = st_read("../data/allPoints/Sizov_pts.shp", "Sizov_pts") 
f14 = st_read("../data/allPoints/Veraverbeke_pts.shp", "Veraverbeke_pts") 
```

# Check columns names
```{r}
colnames(f01)
```

```{r}
colnames(f02)
```
f02 has added columns "fire_d1"  "fr_yr_1"  "fire_d2"  "fr_yr_2"  "fire_d3"  "fr_yr_3" 
```{r}
colnames(f03)
```
```{r}
colnames(f04)

```
```{r}
colnames(f05)
```
```{r}
colnames(f06)
```
```{r}
colnames(f07)
```
```{r}
colnames(f08)
```
```{r}
colnames(f09)
```
```{r}
colnames(f10)
```
f10 is missing "fire_id"
```{r}
colnames(f11)
```

```{r}
colnames(f12)
```
```{r}
colnames(f13)
```
```{r}
colnames(f14)
```


# drop fire_id
Remove fire_id as some are character and some are integer
```{r}
( pts01 = f01 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )

( pts02 = f02 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```

```{r}
( pts03 = f03 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```

```{r}
( pts04 = f04 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )


( pts05 = f05 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```
```{r}
( pts06 = f06 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )


( pts07 = f07 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```

```{r}
( pts08 = f08 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )

( pts09 = f09 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```

```{r}
( pts10 = f10 %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
  
( pts11 = f11 %>% 
  dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```

```{r}
( pts12 = f12 %>% 
 dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )


( pts13 = f13 %>% dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )


( pts14 = f14 %>% dplyr::select(-fire_id) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope, day), as.numeric)) %>% 
    mutate(across(day, ~coalesce(., 0))) )
```



# Combine all shapefiles
```{r}
all_pts = dplyr::bind_rows(pts01, pts02, pts03, pts04, pts05, pts06, pts07, pts08, pts09, pts10, pts11, pts12, pts13, pts14)
```

# Look at combined data
```{r}
head(all_pts)
```

```{r}
tail(all_pts)
```
## Check factor values
```{r}
unique(all_pts$brl_tnd)
```
```{r}
unique(all_pts$thw_ctv)
```
```{r}
min(all_pts$thw_dpt)
max(all_pts$thw_dpt)
```
```{r}
all_pts[is.na(all_pts$thw_ctv),]
all_pts[is.na(all_pts$thw_dpt),]
```
```{r}
all_pts %>% filter(last_nm=="Sizov, Soromotin, Brodt")
```

```{r}
unique(all_pts$brn_nbr)
```

## Clean up factors
```{r}
( all_pts_clean = all_pts %>%
  mutate(thw_ctv = recode_factor(thw_ctv, "T" = "thaw", A = "active")) %>%
  mutate(brl_tnd = recode_factor(brl_tnd, B = "boreal", boreal = "boreal", "T" = "tundra", tundra = "tundra")) %>%
  mutate(brn_nbr = recode_factor(brn_nbr, burned = "burned", burn = "burned", Burned = "burned", Burn = "burned", '72burn' = "burned", Unburn = 'unburned', unburn = 'unburned', unburned = 'unburned')) %>%
    drop_na(thw_ctv) %>%
    drop_na(thw_dpt)
  )
```
## Column Names for EE
cat(paste0(sprintf('"%s"', xdf$Name), collapse = ", "))
```{r}
cat(paste0(sprintf('"%s"', colnames(all_pts_clean)), collapse = ", "))
```
```{r}
( 
  julianDt_na = all_pts_clean %>% filter_at(vars(julinDt), all_vars(is.na(.))) 
)

unique(julianDt_na$last_nm)
```
```{r}
(
  dropJulianDtNA = all_pts_clean %>% filter(julinDt>0)
)
```
## Check factos again

```{r}
unique(all_pts_clean$brl_tnd)
```
```{r}
unique(all_pts_clean$thw_ctv)
```

```{r}
all_pts_clean[is.na(all_pts_clean$thw_ctv),]
```

```{r}
unique(all_pts_clean$brn_nbr)
```
```{r}
unique(all_pts_clean$year)
```
```{r}
min(all_pts_clean$thw_dpt)
max(all_pts_clean$thw_dpt)
drop = all_pts_clean %>% filter(thw_dpt < 100)
hist(drop$thw_dpt)
```

# write to check data
```{r eval=FALSE, include=FALSE}
st_write(all_pts_clean, "../outputs/allPoints/AllDataPts.shp", driver="ESRI Shapefile") 
```

```{r eval=FALSE, include=FALSE}
st_write(dropJulianDtNA, "../outputs/allPoints/AllDataPtsRemoveJulianNA.shp", driver="ESRI Shapefile") 
```

# Create summarized points by site for map

```{r}
unique(all_pts$site_id)
```
```{r}
unique(all_pts$brn_nbr)
```
```{r}
unique(all_pts$brl_tnd)
```

```{r}
unique(all_pts$thw_ctv)
```

new_data <- data %>% filter_all(any_vars(is.na(.))) 
```{r}
all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, brn_nbr, geometry) %>%
  st_set_geometry(NULL) %>% filter_all(any_vars(is.na(.))) 
```

```{r}
all_pts %>% filter(last_nm =="Delcourt-Veraverbeke")
```


```{r}
all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr, brn_nbr, geometry) %>%
  st_set_geometry(NULL) %>%
  mutate(id = row_number()) %>%
  mutate(n=as.integer(1)) %>%
  filter(last_nm =="Delcourt-Veraverbeke") %>%
  pivot_wider( names_from = brn_nbr, values_from = n, id_cols = c(id, last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr)) %>%
  mutate(across(c(burned, unburned), ~coalesce(., 0))) %>%
  group_by(last_nm, brl_tnd, site_id) %>%
  summarize( lat = mean(lat),
             long = mean(long),
             burned = sum(burned),
             unburned = sum(unburned))
```

# Summarize all points
"tundra" "boreal" "T"      "B"  
```{r}
(
  sum_all_pts = all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr, brn_nbr, thw_ctv, geometry) %>%
  st_set_geometry(NULL) %>%
  mutate(id = row_number()) %>%
  mutate(n=as.integer(1)) %>%
  mutate(brl_tnd = recode_factor(brl_tnd, B = "boreal", boreal = "boreal", "T" = "tundra", tundra = "tundra")) %>%
  mutate(brn_nbr = recode_factor(brn_nbr, burned = "burned", burn = "burned", Burned = "burned", Burn = "burned", '72burn' = "burned", Unburn = 'unburned', unburn = 'unburned', unburned = 'unburned')) %>%
  group_by(id) %>%
  pivot_wider( names_from = brn_nbr, values_from = n, id_cols = c(id, last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr)) %>%
  mutate(across(c(burned, unburned), ~coalesce(., 0))) %>%
  group_by(last_nm, brl_tnd) %>%
  summarize( lat = mean(lat),
             long = mean(long),
             burned = sum(burned),
             unburned = sum(unburned)) %>%
  mutate(total = burned + unburned) %>%
  mutate(propBurn = burned/total)
)

 

```

## back to shapefile
```{r}
sum_all_pts_sf = st_as_sf(sum_all_pts, coords = c("long","lat"), crs = 4326, remove = FALSE)
```

# write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(sum_all_pts_sf, "../outputs/allPoints/SumAllPts.shp", driver="ESRI Shapefile")
```

# Aggregate points within 100 km of eachother
```{r}
all_pts_clean
```
```{r}
unique(all_pts_clean$plot_id)
```

```{r}
unique(all_pts_clean$site_id)
```

```{r}
( pts_agg <- all_pts_clean %>% group_by(last_nm) %>%
    aggregate(all_pts_clean,
                     FUN = mean, 
                     join = function(x, y) st_is_within_distance(x, y, dist = 100)) )
```
