---
title: "Combine all point files for mapping"
author: "Anna Talucci"
date: "2023-03-22"
output: html_document
---
# clear environment
```{r}
rm(list=ls())
```

# Overview

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data
```{r}
f01 = st_read("../data/allPoints/Baillargeon_pts.shp", "Baillargeon_pts") 
f02 = st_read("../data/allPoints/Buma_pts.shp", "Buma_pts") 
f03 = st_read("../data/allPoints/Dielman_pts.shp", "Dielman_pts") 
f04 = st_read("../data/allPoints/Douglas_pts.shp", "Douglas_pts") 
f05 = st_read("../data/allPoints/Frost_pts.shp", "Frost_pts") 
f06 = st_read("../data/allPoints/Gaglioti_pts.shp", "Gaglioti_pts") 
f07 = st_read("../data/allPoints/Manies_pts.shp", "Manies_pts") 
f08 = st_read("../data/allPoints/Natali_pts.shp", "Natali_pts") 
f09 = st_read("../data/allPoints/ODonnell_pts.shp", "ODonnell_pts") 
f10 = st_read("../data/allPoints/Olefeldt_pts.shp", "Olefeldt_pts") 
f11 = st_read("../data/allPoints/Paulson_pts.shp", "Paulson_pts")
f12 = st_read("../data/allPoints/Rocha_pts.shp", "Rocha_pts") 
f13 = st_read("../data/allPoints/Sizov_pts.shp", "Sizov_pts") 
f14 = st_read("../data/allPoints/Veraverbeke_pts.shp", "Veraverbeke_pts") 
```

# Check columns names
```{r}
colnames(f01)
```

```{r}
colnames(f02)
```
f02 has added columns "fire_d1"  "fr_yr_1"  "fire_d2"  "fr_yr_2"  "fire_d3"  "fr_yr_3" 
```{r}
colnames(f03)
```
```{r}
colnames(f04)

```
```{r}
colnames(f05)
```
```{r}
colnames(f06)
```
```{r}
colnames(f07)
```
```{r}
colnames(f08)
```
```{r}
colnames(f09)
```
```{r}
colnames(f10)
```
f10 is missing "fire_id"
```{r}
colnames(f11)
```

```{r}
colnames(f12)
```
```{r}
colnames(f13)
```
```{r}
colnames(f14)
```


# drop fire_id
Remove fire_id as some are character and some are integer
```{r}
f01 = f01 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f02 = f02 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
```

```{r}
f03 = f03 %>% 
  dplyr::select(-fire_id, -year, -month, -day) %>% 
  mutate(across(c(plot_id, site_id), as.character)) %>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
```

```{r}
f04 = f04 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f05 = f05 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
```
```{r}
f06 = f06 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f07 = f07 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f08 = f08 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f09 = f09 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f10 = f10 %>% dplyr::select(-year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f11 = f11 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f12 = f12 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f13 = f13 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
f14 = f14 %>% dplyr::select(-fire_id, -year, -month, -day) %>% mutate(across(c(plot_id, site_id), as.character))%>% 
  mutate(across(c(fire_yr, orgnc_d, slope), as.numeric))
```



# Combine all shapefiles
```{r}
all_pts = dplyr::bind_rows(f01, f02, f03, f04, f05, f06, f07, f08, f09, f10, f11, f12, f13, f14)
```

# Look at combined data
```{r}
head(all_pts)
```

```{r}
tail(all_pts)
```

```{r}
all_pts
```



# write to check data
```{r eval=FALSE, include=FALSE}
st_write(all_pts, "../outputs/allPoints/AllDataPts.shp", driver="ESRI Shapefile") 
```

# Create summarized points by site for map

```{r}
unique(all_pts$site_id)
```
```{r}
unique(all_pts$brn_nbr)
```
```{r}
unique(all_pts$brl_tnd)
```
new_data <- data %>% filter_all(any_vars(is.na(.))) 
```{r}
all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, brn_nbr, geometry) %>%
  st_set_geometry(NULL) %>% filter_all(any_vars(is.na(.))) 
```

```{r}
all_pts %>% filter(last_nm =="Delcourt-Veraverbeke")
```


```{r}
all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr, brn_nbr, geometry) %>%
  st_set_geometry(NULL) %>%
  mutate(id = row_number()) %>%
  mutate(n=as.integer(1)) %>%
  filter(last_nm =="Delcourt-Veraverbeke") %>%
  pivot_wider( names_from = brn_nbr, values_from = n, id_cols = c(id, last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr)) %>%
  mutate(across(c(burned, unburned), ~coalesce(., 0))) %>%
  group_by(last_nm, brl_tnd, site_id) %>%
  summarize( lat = mean(lat),
             long = mean(long),
             burned = sum(burned),
             unburned = sum(unburned))
```

# Summarize all points
"tundra" "boreal" "T"      "B"  
```{r}
(
  sum_all_pts = all_pts %>% 
  dplyr::select(last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr, brn_nbr, geometry) %>%
  st_set_geometry(NULL) %>%
  mutate(id = row_number()) %>%
  mutate(n=as.integer(1)) %>%
  mutate(brl_tnd = recode_factor(brl_tnd, B = "boreal", boreal = "boreal", "T" = "tundra", tundra = "tundra")) %>%
  mutate(brn_nbr = recode_factor(brn_nbr, burned = "burned", burn = "burned", Burned = "burned", Burn = "burned", '72burn' = "burned", Unburn = 'unburned', unburn = 'unburned', unburned = 'unburned')) %>%
  group_by(id) %>%
  pivot_wider( names_from = brn_nbr, values_from = n, id_cols = c(id, last_nm, cntry_c, lat, long, brl_tnd, site_id, plot_id, fire_yr)) %>%
  mutate(across(c(burned, unburned), ~coalesce(., 0))) %>%
  group_by(last_nm, brl_tnd, site_id) %>%
  summarize( lat = mean(lat),
             long = mean(long),
             burned = sum(burned),
             unburned = sum(unburned)) %>%
  mutate(total = burned + unburned) %>%
  mutate(propBurn = burned/total)
)

 

```

## back to shapefile
```{r}
sum_all_pts_sf = st_as_sf(sum_all_pts, coords = c("long","lat"), crs = 4326, remove = FALSE)
```

# write to shapefile
```{r}
st_write(sum_all_pts_sf, "../outputs/allPoints/SumAllPts.shp", driver="ESRI Shapefile")
```

