---
title: "Douglas Clean to Point"
author: "Anna Talucci"
date: '2022-12-11'
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Douglas
```{r}
f5 <- read.csv(f[5], header = TRUE)
f5
```
```{r}
str(f5)
```

# fix incorrect logical columns
```{r}
f5$slope <- read.csv(f[5], header = TRUE, colClasses = "character")[,20]
```

```{r}
f5_dates = f5 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) %>%
  dplyr::select(-notes) %>%
  drop_na(.)

f5_dates
```




```{r}
f5_dates %>% distinct(julianDate, year, date)
```

# Split by active & thaw
```{r}
f5_thaw = f5_dates %>% 
  filter(thaw_active == "T") 

f5_thaw

f5_active = f5_dates %>% 
  filter(thaw_active == "A") 

f5_active
```

# Create points shapefile

## For F5
sf::st_as_sf(dd, coords = c("x","y"))
```{r}
f5_active_pts = st_as_sf(f5_active, coords = c("long","lat"), crs = 4326, remove = FALSE)
```

```{r}
f5_pts = st_as_sf(f5_dates, coords = c("long","lat"), crs = 4326, remove = FALSE)
```


### write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(f5_active_pts, "../outputs/ActivePoints/Douglas_active_pts.shp", driver="ESRI Shapefile")
```

```{r}
st_write(f5_pts, "../outputs/allPoints/Douglas_pts.shp", driver="ESRI Shapefile")
```