---
title: "ERA5 Data"
author: "Anna Talucci"
date: '2022-12-02'
output: html_document
---

# Overview

Daily Aggregates https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_DAILY

Mean daily air temperature converted from Kelvin to Celsius in earth engine

# Packages
```{r}
library(tidyverse)
library(lubridate)
```

# Data 

## Air Temperature
```{r}
era5 = read.csv("../data/ERA5LandT2M_TEST.csv")
```
```{r}
era5_mean = read.csv("../data/ERA5DailyMeanT2M_TEST.csv")

```


```{r}
era5_v3 = read.csv("../data/ERA5LandT2M_TEST3.csv")
```

```{r}
head(era5)
```
```{r}
era5_v3
```
```{r}
str(era5_v3)
```
```{r}
names(era5_v3)
```

```{r}
scale2 <- function(x, na.rm = FALSE) (x - 273.15)
```

```{r}
replace2 = function(x) (na_if(x<0))
```

```{r}
era5_v3 %>% 
  dplyr::select(-system.index, -.geo) %>%
  mutate_at(vars(matches("temperature_2m")), scale2) %>% 
  mutate_at(vars(matches("temperature_2m")), ~ifelse(.<=0,"NA",.)) 
```
Drop first and last column, scale temperature from Kelvin to Celcius, make negative temperatures zero, sum temperature by row.

New data is already converted to Celsius
```{r}
unique(era5_mean$julinDt)
```
```{r}
head(era5_mean)
```
# 211
Julian Data 211 2019-07-30
```{r}
jd211 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==211) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190730_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd212 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==212) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190731_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd213 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==213) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190801_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```


```{r}
jd214 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==214) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190802_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd216 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==216) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190804_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd217 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==217) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190805_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd218 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==218) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190806_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

```{r}
jd219 = era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==219) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("mean_2m_air_temperature")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(A= sum(across(c(X20190401_mean_2m_air_temperature:X20190807_mean_2m_air_temperature)))) %>%
  mutate(B = sum(across(c(X20190401_mean_2m_air_temperature:X20190929_mean_2m_air_temperature)))) %>%
  mutate(sqrtA = sqrt(A)) %>%
  mutate(sqrtB = sqrt(B)) %>%
  mutate(C = B/A) %>%
  mutate(adjALT = (thw_dpt*C))
  
```

# Combine
```{r}
adjustedData = bind_rows(list(jd211, jd212, jd213, jd214, jd216, jd217, jd218, jd219))
```

```{r}
adjALTData = adjustedData %>% 
  dplyr::select(brn_nbr:cntry_c, day:adjALT)
```

# Save new data
```{r}
write_csv(adjALTData, "../outputs/Veraverbeke_AdjAlt.csv")
```

## Older versions
```{r}
era5_mean %>% 
  dplyr::select(-system.index, -.geo) %>%
  filter(julinDt==211) %>%
  #Subset temp columns specifc to julian dates of interest 
  #filter by julian date 
  mutate_at(vars(matches("temperature_2m")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(TDD = sum(across(ends_with("temperature_2m"))))
```

```{r}
era5_v3 %>% 
  dplyr::select(-system.index, -.geo) %>%
  mutate_at(vars(matches("temperature_2m")), scale2) %>% 
  mutate_at(vars(matches("temperature_2m")), ~ifelse(.<=0, 0,.)) %>% 
  rowwise() %>% 
    mutate(rowsum = sum(across(ends_with("temperature_2m"))))
```

mutate(sumrow = ifelse(julinDt == 211, "D",
                                     ifelse(grade %in% 70:79, "C",
                                            ifelse(grade %in% 80:89, "B",
                                                   ifelse(grade %in% 90:99, "A", "F")))))

```{r}
era5_v3 %>% 
  dplyr::select(-system.index, -.geo) %>%
  mutate_at(vars(matches("temperature_2m")), scale2) %>% 
  rowwise() %>% 
    mutate(rowsum = sum(across(ends_with("temperature_2m"))))
  
```

```{r}
era5_v3$X105_temperature_2m
```