---
title: "ERA5 Data Organization"
author: "Anna Talucci"
date: "2023-03-25"
output: html_document
---



# clear environment
```{r}
rm(list=ls())
```

# Overview

ERA 5 data from google earth engine


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
library(data.table)
```



# Data

## vector of data files
```{r}
f <- list.files(path = "../data/ERA5DataEE/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Individual files
```{r}
f2001 <- read.csv(f[1], header = TRUE) 
f2005 <- read.csv(f[2], header = TRUE) 
f2006 <- read.csv(f[3], header = TRUE) 
f2007 <- read.csv(f[4], header = TRUE) 
f2008 <- read.csv(f[5], header = TRUE) 
f2009 <- read.csv(f[6], header = TRUE) 
f2010 <- read.csv(f[7], header = TRUE) 
f2011 <- read.csv(f[8], header = TRUE) 
f2012 <- read.csv(f[9], header = TRUE) 
f2013 <- read.csv(f[10], header = TRUE) 
f2014 <- read.csv(f[11], header = TRUE) 
f2015 <- read.csv(f[12], header = TRUE) 
f2016 <- read.csv(f[13], header = TRUE) 
f2017 <- read.csv(f[14], header = TRUE) 
f2018 <- read.csv(f[15], header = TRUE) 
f2019 <- read.csv(f[16], header = TRUE) 
f2020 <- read.csv(f[17], header = TRUE) 
```

## View DF
```{r}
f2001
```


## Process
1. pivot longer so column of temperature and date
2. Group by site_id
3. Select dates after day of thaw measurement
4. Select only rows with temp above 0 (zero)
5.  apply calculation

```{r}
( df2001 = f2001 %>%
  dplyr::select(brl_tnd:year, X20010401_temperature_2m:X20011031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2005 = f2005 %>%
  dplyr::select(brl_tnd:year, X20050401_temperature_2m:X20051031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2006 = f2006 %>%
  dplyr::select(brl_tnd:year, X20060401_temperature_2m:X20061031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2007 = f2007 %>%
  dplyr::select(brl_tnd:year, X20070401_temperature_2m:X20071031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2008 = f2008 %>%
  dplyr::select(brl_tnd:year, X20080401_temperature_2m:X20081031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2009 = f2009 %>%
  dplyr::select(brl_tnd:year, X20090401_temperature_2m:X20091031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2010 = f2010 %>%
  dplyr::select(brl_tnd:year, X20100401_temperature_2m:X20101031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2011 = f2011 %>%
  dplyr::select(brl_tnd:year, X20110401_temperature_2m:X20111031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2012 = f2012 %>%
  dplyr::select(brl_tnd:year, X20120401_temperature_2m:X20121031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2013 = f2013 %>%
  dplyr::select(brl_tnd:year, X20130401_temperature_2m:X20131031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2014 = f2014 %>%
  dplyr::select(brl_tnd:year, X20140401_temperature_2m:X20141031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2015 = f2015 %>%
  dplyr::select(brl_tnd:year, X20150401_temperature_2m:X20151031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2016 = f2016 %>%
  dplyr::select(brl_tnd:year, X20160401_temperature_2m:X20161031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2017 = f2017 %>%
  dplyr::select(brl_tnd:year, X20170401_temperature_2m:X20171031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2018 = f2018 %>%
  dplyr::select(brl_tnd:year, X20180401_temperature_2m:X20181031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
( df2019 = f2019 %>%
  dplyr::select(brl_tnd:year, X20190401_temperature_2m:X20191031_temperature_2m) %>%
    mutate(id = row_number()) )
```

```{r}
f2020
( df2020 = f2020 %>%
  dplyr::select(brl_tnd:year, X20200401_temperature_2m:X20201031_temperature_2m) %>%
    mutate(id = row_number()) )
```

# Functions to clean 
```{r}
cleanDf = function(x)  {
  x %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(plot_id = as.character(plot_id)) %>%
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) %>%
    mutate(tempDate=ymd(tempDate)) %>%
    mutate(julianTemp = yday(tempDate)) -> x
  return(x)
}
```


# Apply functions to to each year
## Clean function
```{r}
( clean2001 = df2001 %>% cleanDf() )
```


```{r}
clean2005 = df2005 %>% cleanDf()
clean2006 = df2006 %>% cleanDf()
clean2007 = df2007 %>% cleanDf()
clean2008 = df2008 %>% cleanDf()
clean2009 = df2009 %>% cleanDf()
```
FYI the warning is referencing parts of the name being discarded in the separate process. This can be ignored for these chunks of code.
```{r}
clean2010 = df2010 %>% cleanDf()
clean2011 = df2011 %>% cleanDf()
clean2012 = df2012 %>% cleanDf()
clean2013 = df2013 %>% cleanDf()
clean2014= df2014 %>% cleanDf()
```

```{r}
clean2015 = df2015 %>% cleanDf()
clean2016 = df2016 %>% cleanDf()
clean2017 = df2017 %>% cleanDf()
clean2018 = df2018 %>% cleanDf()
clean2019 = df2019 %>% cleanDf()
clean2020 = df2020 %>% cleanDf()
```

# write to csv
```{r eval=FALSE, include=FALSE}
write.csv(clean2001,file = "../outputs/era5Clean/clean2001.csv")
write.csv(clean2005,file = "../outputs/era5Clean/clean2005.csv")
write.csv(clean2006,file = "../outputs/era5Clean/clean2006.csv")
write.csv(clean2007,file = "../outputs/era5Clean/clean2007.csv")
write.csv(clean2008,file = "../outputs/era5Clean/clean2008.csv")
```

```{r eval=FALSE, include=FALSE}
write.csv(clean2009,file = "../outputs/era5Clean/clean2009.csv")
write.csv(clean2010,file = "../outputs/era5Clean/clean2010.csv")
write.csv(clean2011,file = "../outputs/era5Clean/clean2011.csv")
write.csv(clean2012,file = "../outputs/era5Clean/clean2012.csv")
write.csv(clean2013,file = "../outputs/era5Clean/clean2013.csv")
```

```{r eval=FALSE, include=FALSE}
write.csv(clean2014,file = "../outputs/era5Clean/clean2014.csv")
write.csv(clean2015,file = "../outputs/era5Clean/clean2015.csv")
write.csv(clean2016,file = "../outputs/era5Clean/clean2016.csv")
write.csv(clean2017,file = "../outputs/era5Clean/clean2017.csv")
write.csv(clean2018,file = "../outputs/era5Clean/clean2018.csv")
```

```{r eval=FALSE, include=FALSE}
write.csv(clean2019,file = "../outputs/era5Clean/clean2019.csv")
write.csv(clean2020,file = "../outputs/era5Clean/clean2020.csv")
```

## Identify ALT date

### 2001
```{r}
( rm2001 = clean2001 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp>=200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

( distinct2001 = rm2001 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```

```{r}
( act = clean2001 %>%
  filter(thw_ctv =="active") 
)

unique(act$julinDt)
```

### 2005
```{r}
( rm2005 = clean2005 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp>=200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

(distinct2005 = rm2005 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2006 
**2006 is active only**
```{r}
clean2006
( rm2006 = clean2006 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 220)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

rm2006 %>% distinct(site_id, Goal2, Goal3)

```
### 2007
**2007 is active only**
```{r}

( rm2007 = clean2007 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 220)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

rm2007 %>% distinct(site_id, Goal2, Goal3)

```
### 2008
```{r}

( rm2008 = clean2008 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

(distinct2008 = rm2008 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )
```
### 2009
```{r}

( rm2009 = clean2009 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

( distinct2009 = rm2009 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )
```
### 2010
```{r}

( rm2010 = clean2010 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

( distinct2010 = rm2010 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )
```
### 2011
```{r}

( rm2011 = clean2011 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )

( distinct2011 = rm2011 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )
```
### 2012
```{r}

( rm2012 = clean2012 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2012 = rm2012 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2013
```{r}

( rm2013 = clean2013 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2013 = rm2013 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2014
```{r}

( rm2014 = clean2014 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2014 = rm2014 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2015
```{r}

( rm2015 = clean2015 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2015 = rm2015 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2016
```{r}

( rm2016 = clean2016 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2016 = rm2016 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2017
```{r}

( rm2017 = clean2017 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2017 = rm2017 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2018
```{r}

( rm2018 = clean2018 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2018 = rm2018 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
### 2019
```{r}

( rm2019 = clean2019 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2019 = rm2019 %>% distinct(plot_id, site_id, year,julianTemp, Goal2, Goal3) )

```
### 2020
```{r}

( rm2020 = clean2020 %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(thw_ctv =="thaw") %>% 
    filter(julianTemp >= 200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(last_nm, plot_id, site_id) %>%
    arrange(julianTemp) %>%
  mutate(Goal = cumsum(c(1, diff(tempDate) > 1))) %>%
  group_by(plot_id, site_id, Goal) %>%
  mutate(Goal2 = min(tempDate), Goal3 = max(tempDate))
    
  )


( distinct2020 = rm2020 %>% distinct(plot_id, site_id, year, julianTemp, Goal2, Goal3) )

```
# Bind clean thaw
```{r}
( 
rm = dplyr::bind_rows(rm2001, rm2005, rm2008, rm2009, rm2010, rm2011, rm2012, rm2013, rm2014, rm2015, rm2016, rm2017, rm2018, rm2019, rm2020)
)
```


```{r}

```


# Bind distinct df
```{r}
( 
distinct = dplyr::bind_rows(distinct2001, distinct2005, distinct2008, distinct2009, distinct2010, distinct2011, distinct2012, distinct2013, distinct2014, distinct2015, distinct2016, distinct2017, distinct2018, distinct2019, distinct2020)
)
```

```{r}
( 
  summarizeDistinct = distinct %>%
  group_by(plot_id, site_id, year) %>%
  summarise( mean_Julian = mean(julianTemp),
             mean_goal2 = mean(Goal2),
             min_goal2 = min(Goal2),
             max_goal2 = max(Goal2)) %>%
  arrange(min_goal2) 
  
)
  
```

```{r}
summarizeDistinct %>% 
  group_by(plot_id, site_id) %>%
  mutate(KEEP = any(min_goal2 == max_goal2)) %>% 
  filter(KEEP == FALSE)
```

```{r}
unique(distinct$plot_id) 
```

# add standard active layer thickness date column
```{r}
( StdDate2001 = clean2001 %>%
  mutate(stdDateAlt = 282-14)
)
```

Make dataframe of distinct potential dates for standardizing by plot, site, name, and year

```{r}
StdDate2005 = clean2005 %>% mutate(stdDateAlt = 273-14)
StdDate2008 = clean2008 %>% mutate(stdDateAlt = 253-14)
StdDate2009 = clean2009 %>% mutate(stdDateAlt = 262-14)
StdDate2010 = clean2010 %>% mutate(stdDateAlt = 265-14)
StdDate2011 = clean2011 %>% mutate(stdDateAlt = 302-14)
```

**What if the standardization date is before the measure date??**
```{r}
StdDate2012 = clean2012 %>% mutate(stdDateAlt = 253-14)
```

```{r}

```


