---
title: "Explore Air Temp data"
author: "Anna Talucci"
date: '2023-01-26'
output: html_document
---

# Overview

# Packages
```{r}
library(tidyverse)
library(lubridate)
library(cowplot)
```

# Graph theme
```{r}

reg_theme = theme_bw() + theme(legend.position = "none") +
  theme(plot.margin = unit(c(t = 0.3, r = 0.3, b = 0.3, l = 0.1), "cm")) +
  theme(axis.title.x = element_text(size = 11, hjust = 0.5, vjust = -0.1),
        axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

```

# Data
```{r}
manies = read.csv("../data/ERA5forThawPoints/ERA5DailyMeanT2M_ManiesSept30.csv")
natali2017 = read.csv("../data/ERA5forThawPoints/2017ERA5DailyMeanT2M_NataliSept30.csv")
natali2016 = read.csv("../data/ERA5forThawPoints/2016ERA5DailyMeanT2M_NataliSept30.csv")
odonnell2005 = read.csv("../data/ERA5forThawPoints/2005ERA5DailyMeanT2M_ODonnellSept30.csv")
olefeldt2015 = read.csv("../data/ERA5forThawPoints/2015ERA5DailyMeanT2M_OlefeldtSept30.csv")
olefeldt2016 = read.csv("../data/ERA5forThawPoints/2016ERA5DailyMeanT2M_OlefeldtSept30.csv")
olefeldt2020 = read.csv("../data/ERA5forThawPoints/2020ERA5DailyMeanT2M_OlefeldtSept30.csv")
Rocha2008 = read.csv("../data/ERA5forThawPoints/2008ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2009 = read.csv("../data/ERA5forThawPoints/2009ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2010 = read.csv("../data/ERA5forThawPoints/2010ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2011 = read.csv("../data/ERA5forThawPoints/2011ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2012 = read.csv("../data/ERA5forThawPoints/2012ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2013 = read.csv("../data/ERA5forThawPoints/2013ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2014 = read.csv("../data/ERA5forThawPoints/2014ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2015 = read.csv("../data/ERA5forThawPoints/2015ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2016 = read.csv("../data/ERA5forThawPoints/2016ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2017 = read.csv("../data/ERA5forThawPoints/2017ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2018 = read.csv("../data/ERA5forThawPoints/2018ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2019 = read.csv("../data/ERA5forThawPoints/2019ERA5DailyMeanT2M_RochaSept30.csv")
Rocha2020 = read.csv("../data/ERA5forThawPoints/2020ERA5DailyMeanT2M_RochaSept30.csv")
```

# Look at data
```{r}
head(manies)
```
# Organize and Clean

## Manies
### Select, Pivot Air Temp data
```{r}
manies1 = manies %>%
  dplyr::select(brl_tnd:year, X20010401_mean_2m_air_temperature:X20010930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
manies1$tempDate = ymd(manies1$tempDate)
```

### Add Julian day for air temp
```{r}
manies2 = manies1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
manies2
```
### Plot Temp data by day

243 267 215 266 218
```{r}
p1 = ggplot(manies2, aes(x=tempJuldt, y=TempC)) +
  geom_line() +
  geom_vline(xintercept=215, colour = "red") + # Aug 3
  geom_vline(xintercept=267, colour="red") + #Sept 24
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=215, y=-5,label="Aug 3"), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=267, y=-5, label="Sept 24"), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Manies 2005",), colour="grey",  text=element_text(size=7)) 

p1
```

## Natali 2016
### Select, Pivot Air Temp data
```{r}
natali2016_1 = natali2016 %>%
  dplyr::select(brl_tnd:year, X20160401_mean_2m_air_temperature:X20160930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
natali2016_1$tempDate = ymd(natali2016_1$tempDate)
```

### Add Julian day for air temp
```{r}
natali2016_2 = natali2016_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
natali2016_2
```

### Plot Temp data by day

Aug 2
Sept 6
Sept 15, 16, 17, 18, 19

249 214 260  NA 261 259 262
```{r}
p2 = ggplot(natali2016_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=214, colour = "red") + # Aug 2
  geom_vline(xintercept=262, colour = "red") + #Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=214, label="Aug 2", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=259, label="Sept 15", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 19, label="Natali 2016"), colour="grey", text=element_text(size=7)) 
```

## Natali 2017
### Select, Pivot Air Temp data
```{r}
natali2017_1 = natali2017 %>%
  dplyr::select(brl_tnd:year, X20170401_mean_2m_air_temperature:X20170930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
natali2017_1$tempDate = ymd(natali2017_1$tempDate)
```

### Add Julian day for air temp
```{r}
natali2017_2 = natali2017_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
natali2017_2
```

### Plot Temp data by day

Aug 2
Sept 6
Sept 15, 16, 17, 18, 19

249 214 260  NA 261 259 262
```{r}
p3 = ggplot(natali2017_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=214, colour = "red") + # Aug 2
  geom_vline(xintercept=262, colour = "red") + #Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=214, label="Aug 2", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=259, label="Sept 15", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 20, label="Natali 2017"), colour="grey", text=element_text(size=7)) 
```

## O'Donnell 2005
### Select, Pivot Air Temp data
```{r}
odonnell2005_1 = odonnell2005 %>%
  dplyr::select(brl_tnd:year, X20050401_mean_2m_air_temperature:X20050930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
odonnell2005_1$tempDate = ymd(odonnell2005_1$tempDate)
```

### Add Julian day for air temp
```{r}
odonnell2005_2 = odonnell2005_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
odonnell2005_2
```

### Plot Temp data by day

"2007-09-11" "2007-09-12" "2007-09-14" "2007-09-13" "2007-09-15" "2007-09-17" "2008-09-03"
 "2008-09-04" "2008-09-05"

254 255 257 256 258 260 247 248 249
```{r}
p4 = ggplot(odonnell2005_2, aes(x=tempJuldt, y=TempC)) +
  geom_line() +
  geom_vline(xintercept=247, colour = "red") + # Sept 3
  geom_vline(xintercept=260, colour = "red") + # Sept 17
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature  (Celecius)") +
  geom_text(aes(x=247, label="Sept 3", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=260, label="Sept 17", y=-5), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=175, y= -10, label="O'Donnell 2005"), colour="grey",  text=element_text(size=7)) 
```

## Olefeldt 2015
### Select, Pivot Air Temp data
```{r}
olefeldt2015_1 = olefeldt2015 %>%
  dplyr::select(brl_tnd:year, X20150401_mean_2m_air_temperature:X20150930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
olefeldt2015_1$tempDate = ymd(olefeldt2015_1$tempDate)
```

### Add Julian day for air temp
```{r}
olefeldt2015_2 = olefeldt2015_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
olefeldt2015_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p5 = ggplot(olefeldt2015_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour = "red") + # Sept 8
  geom_vline(xintercept=261, colour = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=175, y= -10, label="Olefeldt 2015"), colour="grey", text=element_text(size=7)) 
```

## Olefeldt 2016
### Select, Pivot Air Temp data
```{r}
olefeldt2016_1 = olefeldt2016 %>%
  dplyr::select(brl_tnd:year, X20160401_mean_2m_air_temperature:X20160930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
olefeldt2016_1$tempDate = ymd(olefeldt2016_1$tempDate)
```

### Add Julian day for air temp
```{r}
olefeldt2016_2 = olefeldt2016_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
olefeldt2016_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p6 = ggplot(olefeldt2016_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour = "red") + # Sept 8
  geom_vline(xintercept=261, colour = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=175, y= -10, label="Olefeldt 2016"), colour="grey", text=element_text(size=7)) 
```

## Olefeldt 2020
```{r}
names(olefeldt2020)
```
### Select, Pivot Air Temp data
```{r eval=FALSE, include=FALSE}
olefeldt2020_1 = olefeldt2020 %>%
  dplyr::select(brl_tnd:year, X20200401_mean_2m_air_temperature:X20200930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r eval=FALSE, include=FALSE}
olefeldt2020_1$tempDate = ymd(olefeldt2020_1$tempDate)
```

### Add Julian day for air temp
```{r eval=FALSE, include=FALSE}
olefeldt2020_2 = olefeldt2020_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r eval=FALSE, include=FALSE}
olefeldt2020_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r eval=FALSE, include=FALSE}
p7 = ggplot(olefeldt2020_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour = "red") + # Sept 8
  geom_vline(xintercept=261, colour = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Olefeldt 2020"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2008
### Select, Pivot Air Temp data
```{r}
Rocha2008_1 = Rocha2008 %>%
  dplyr::select(brl_tnd:year, X20080401_mean_2m_air_temperature:X20080930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2008_1$tempDate = ymd(Rocha2008_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2008_2 = Rocha2008_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2008_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p8 = ggplot(Rocha2008_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour= "red") + # Sept 8
  geom_vline(xintercept=261, colour = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-12), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-12), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2008"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2009
### Select, Pivot Air Temp data
```{r}
Rocha2009_1 = Rocha2009 %>%
  dplyr::select(brl_tnd:year, X20090401_mean_2m_air_temperature:X20090930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2009_1$tempDate = ymd(Rocha2009_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2009_2 = Rocha2009_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2009_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p9 = ggplot(Rocha2009_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour = "red") + # Sept 8
  geom_vline(xintercept=261, color = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-12), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-12), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2009"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2010
### Select, Pivot Air Temp data
```{r}
Rocha2010_1 = Rocha2010 %>%
  dplyr::select(brl_tnd:year, X20100401_mean_2m_air_temperature:X20100930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2010_1$tempDate = ymd(Rocha2010_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2010_2 = Rocha2010_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2010_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p10 = ggplot(Rocha2010_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-11), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-11), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2010"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2011
### Select, Pivot Air Temp data
```{r}
Rocha2011_1 = Rocha2011 %>%
  dplyr::select(brl_tnd:year, X20110401_mean_2m_air_temperature:X20110930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2011_1$tempDate = ymd(Rocha2011_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2011_2 = Rocha2011_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2011_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p11 = ggplot(Rocha2011_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-12), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-12), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2011"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2012
### Select, Pivot Air Temp data
```{r}
Rocha2012_1 = Rocha2012 %>%
  dplyr::select(brl_tnd:year, X20120401_mean_2m_air_temperature:X20120930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2012_1$tempDate = ymd(Rocha2012_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2012_2 = Rocha2012_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2012_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p12 = ggplot(Rocha2012_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-12), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-12), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 14, label="Rocha 2012"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2013
### Select, Pivot Air Temp data
```{r}
Rocha2013_1 = Rocha2013 %>%
  dplyr::select(brl_tnd:year, X20130401_mean_2m_air_temperature:X20130930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2013_1$tempDate = ymd(Rocha2013_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2013_2 = Rocha2013_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2013_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p13 = ggplot(Rocha2013_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-20), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-20), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2013"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2014
### Select, Pivot Air Temp data
```{r}
Rocha2014_1 = Rocha2014 %>%
  dplyr::select(brl_tnd:year, X20140401_mean_2m_air_temperature:X20140930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2014_1$tempDate = ymd(Rocha2014_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2014_2 = Rocha2014_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2014_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p14 = ggplot(Rocha2014_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 12, label="Rocha 2014"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2015
### Select, Pivot Air Temp data
```{r}
Rocha2015_1 = Rocha2015 %>%
  dplyr::select(brl_tnd:year, X20150401_mean_2m_air_temperature:X20150930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2015_1$tempDate = ymd(Rocha2015_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2015_2 = Rocha2015_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2015_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p15 = ggplot(Rocha2015_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour = "red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-15), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-15), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 17, label="Rocha 2015"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2016
### Select, Pivot Air Temp data
```{r}
Rocha2016_1 = Rocha2016 %>%
  dplyr::select(brl_tnd:year, X20160401_mean_2m_air_temperature:X20160930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2016_1$tempDate = ymd(Rocha2016_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2016_2 = Rocha2016_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2016_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p16 = ggplot(Rocha2016_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2016"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2017
### Select, Pivot Air Temp data
```{r}
Rocha2017_1 = Rocha2017 %>%
  dplyr::select(brl_tnd:year, X20170401_mean_2m_air_temperature:X20170930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2017_1$tempDate = ymd(Rocha2017_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2017_2 = Rocha2017_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2017_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p17 = ggplot(Rocha2017_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2017"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2018
### Select, Pivot Air Temp data
```{r}
Rocha2018_1 = Rocha2018 %>%
  dplyr::select(brl_tnd:year, X20180401_mean_2m_air_temperature:X20180930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2018_1$tempDate = ymd(Rocha2018_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2018_2 = Rocha2018_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2018_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p18 = ggplot(Rocha2018_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2018"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2019
### Select, Pivot Air Temp data
```{r}
Rocha2019_1 = Rocha2019 %>%
  dplyr::select(brl_tnd:year, X20190401_mean_2m_air_temperature:X20190930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r}
Rocha2019_1$tempDate = ymd(Rocha2019_1$tempDate)
```

### Add Julian day for air temp
```{r}
Rocha2019_2 = Rocha2019_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r}
Rocha2019_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r}
p19 = ggplot(Rocha2019_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-10), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-10), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2019"), colour="grey", text=element_text(size=7)) 
```

## Rocha 2020
### Select, Pivot Air Temp data
```{r eval=FALSE, include=FALSE}
Rocha2020_1 = Rocha2020 %>%
  dplyr::select(brl_tnd:year, X20200401_mean_2m_air_temperature:X20200930_mean_2m_air_temperature) %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) 
```
### Format air temp date
```{r eval=FALSE, include=FALSE}
Rocha2020_1$tempDate = ymd(Rocha2020_1$tempDate)
```

### Add Julian day for air temp
```{r eval=FALSE, include=FALSE}
Rocha2020_2 = Rocha2020_1 %>% mutate(tempJuldt = yday(tempDate))
```

```{r eval=FALSE, include=FALSE}
Rocha2020_2
```

### Plot Temp data by day

sept 8
Sept 18
```{r eval=FALSE, include=FALSE}
p20  = ggplot(Rocha2020_2, aes(x=tempJuldt, y=TempC, color=site_id)) +
  geom_line() +
  geom_vline(xintercept=251, colour="red") + # Sept 8
  geom_vline(xintercept=261, colour="red") + # Sept 18
  reg_theme +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=251, label="Sept 8", y=-5), colour="grey", angle=90, vjust = -0.5, text=element_text(size=6)) +
  geom_text(aes(x=261, label="Sept 18", y=-5), colour="grey", angle=90, vjust = 0.5, text=element_text(size=6)) + 
  geom_text(aes(x=120, y= 15, label="Rocha 2020"), colour="grey", text=element_text(size=7)) 
```

# Combine plots

```{r fig.height=7, fig.width=6}
g1 = cowplot::plot_grid(p1, p2, p3, p4, p5, p6, ncol = 2)
g1
```
```{r fig.height=7, fig.width=6}
g2 = cowplot::plot_grid(p8, p9, p10, p11, p12, p13, ncol = 2)
g2
```

```{r fig.height=7, fig.width=6}
g3 = cowplot::plot_grid(p14, p15, p16, p17, p18, p19, ncol = 2)
g3
```




```{r eval=FALSE, include=FALSE}
ggsave("../figures/TempDay1.tiff", plot = g1, width = 6, height =7, units = c("in"), dpi=600, bg = "white" )
ggsave("../figures/TempDay2.tiff", plot = g2, width = 6, height =7, units = c("in"), dpi=600, bg = "white" )
ggsave("../figures/TempDay3.tiff", plot = g3, width = 6, height =7, units = c("in"), dpi=600, bg = "white" )
```