---
title: "Figure Measured Predicted Air Temperature"
author: "Anna Talucci"
date: "2023-06-02"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview
Figures looking at measured versus predicted


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggnewscale)
```
# Plot theme
```{r}
comp_theme = theme_light() + theme(legend.position = "none") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))
```

# Data

## Temp
```{r}
temp = read_csv("../data/era5Clean/ERA5CombineAll.csv")
```

## Standardized measures
```{r}
allDataALT = read_csv("../data/Standardize/AllDataStandardizeALTwithDates.csv")
```


# Filter

## Select first 
```{r}
allDataALT
```
```{r}
( example1 = allDataALT[1, ] )
```

```{r}
( sample150 = allDataALT %>% 
    filter(last_nm == "Manies") %>%
    filter(id ==150) %>%
    rename(GraphJulian = altDate)
)


```

```{r}
( sample151 = allDataALT %>% 
    filter(last_nm == "Manies") %>%
    filter(id == 151) %>%
    rename(GraphJulian = altDate)
)


```

```{r}
( Temp150 = temp %>% 
    filter(last_nm == "Manies") %>%
    filter(id == 150) %>%
    rename(GraphJulian = altDate)
)


```

```{r}
( Temp151 = temp %>% 
    filter(last_nm == "Manies") %>%
    filter(id == 151) %>%
    rename(GraphJulian = altDate) %>%
    mutate(measureALT = 38 ) %>%
    mutate(measureDate = 151 ) %>%
    mutate(predALT = 132 ) %>%
    mutate(predDate = 270 )
)
```

```{r}
( Temp2 = temp %>% 
    filter(TempC > -9) %>% 
    filter(julianTemp > 100) %>%
    filter(julianTemp <290) %>%
    filter(last_nm == "Manies") %>%
    filter(id %in% c(150, 151)) 
)


```


## 2 Active
```{r}
type <- c("measured", "predicted")
depth <- c(-3.68, -3.68)
dayJ <- c(264, 271) # 267, 270, adjust to minimize overlap on graph

df <- data.frame(type, depth, dayJ)
```

position=position_jitter(h=0,w=3),

theme(axis.title = element_text( angle = 90, color="red", size=15, face=3))
```{r}
( plot150Active2 = Temp2 %>%
    filter(id=="150") %>%
    ggplot(., mapping = aes(x = julianTemp, y = TempC)) + 
    geom_line(color='#006D5B')+
   scale_y_continuous(
      name = (expression("Temperature " ( degree*C))), breaks = c(-10,-5, 0, 10, 20),
      sec.axis = dup_axis( name="", breaks=c(0,-1.6, -3.68), labels = c("0","5", "40"))) +
  geom_segment(aes(x = 264, y = 0, xend = 264, yend = -3.68), color="black", size=1) +
  geom_segment(aes(x = 271, y = 0, xend = 271, yend = -3.68), color="black", size=1) +
    # Point1
    new_scale_color() +
    geom_point(data = df, aes(x=dayJ, y=depth, fill=type, color=type),  size = 4, pch=21, stroke =1.8) +# 40
          scale_fill_manual(name="", values=c( "#3b5a9d", "#FEC287FF"))  +
          scale_color_manual(name="", values = c( "grey79", "black")) +

    
    xlab("Day of year") +
    
  theme_light() +
    theme(legend.position = "none", plot.margin = margin(t=5, b=0, r=20, l=5, "mm")) +  

    theme(axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9, color = "black"),
        axis.text.y = element_text(size = 9, color = "black"))
)
```
axis.title.y=element_text(angle=0, face="bold", hjust=0.2,
                                  margin = margin(r = -70))) +

```{r}
( plot150Active = Temp2 %>%
    filter(id=="150") %>%
    ggplot(., mapping = aes(x = julianTemp, y = TempC)) + 
    geom_line(color='#006D5B')+
   scale_y_continuous(
      name = "Temperature (Celecius)", breaks = c(-10,-5, 0, 10, 20),
      sec.axis = dup_axis( name="Depth (cm)\n(log scale)", breaks=c(0,-1.6, -3.68), labels = c("0","-5", "-40"))) +

    geom_point(aes(x=267, y=-3.68), color='#4F7959', size = 2) +# 40
  geom_point(aes(x=270, y=-3.68), color='#4F7959', size = 2) +# 40
  geom_segment(aes(x = 267, y = 0, xend = 267, yend = -3.68), color="#4F7959") +
  geom_segment(aes(x = 270, y = 0, xend = 270, yend = -3.68), color="#4F7959") +
    
    annotate('text', x=262, y=-2.5, label = "Measured", angle=90) +
    annotate('text', x=275, y=-2.5, label = "Predicted", angle=-90) +
    xlab("Julian date") +
  theme_light()
)
```
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  
  theme(axis.title.y.right = element_text(vjust=0.9)) +
```{r}
( plot151Thaw = Temp2 %>%
    filter(id=="150") %>%ggplot(Temp2, mapping = aes(x = julianTemp, y = TempC)) + 
    geom_line(color='#5c7070')+
   scale_y_continuous(
      name = "Temperature (Celecius)",
      sec.axis = dup_axis( name="Depth (cm)\n(log scale)", breaks=c(0,-1.6,-3.46, -4.8), labels = c("0","-5", "-38", "-132"))) +
  geom_point(aes(x=270, y=-4.8), color='#68423D', size=2) + # 132
  geom_point(aes(x=150, y=-3.46), color='#68423D', size=2) +# 38
  geom_segment(aes(x = 150, y = 0, xend = 150, yend = -3.46), color='#68423D') +
  geom_segment(aes(x = 270, y = 0, xend = 270, yend = -4.8), color='#68423D') +
    annotate('text', x=145, y=-2, label = "Measured", angle=90)+
    annotate('text', x=275, y=-2.5, label = "Predicted", angle=-90)+
 xlab("Julian date") +
    
  theme_light()
)
```
## 2 Thaw
```{r}
type2 <- c("measured", "predicted")
depth2 <- c(-3.46, -4.8)
dayJ2 <- c(150, 270) # 267, 270, adjust to minimize overlap on graph

df2 <- data.frame(type2, depth2, dayJ2)
```
```{r}
( plot151Thaw2 = Temp2 %>%
    filter(id=="151") %>%
    ggplot(., mapping = aes(x = julianTemp, y = TempC)) + 
    geom_line(color='#006D5B')+
   scale_y_continuous(
      name = (expression("Temperature " ( degree*C))), breaks = c(-10,-5, 0, 10, 20),
      sec.axis = dup_axis( name="", breaks=c(0,-1.6,-3.46, -4.8), labels = c("0","5", "38", "132"))) +
  geom_segment(aes(x = 150, y = 0, xend = 150, yend = -3.46), color="black", size=1) +
  geom_segment(aes(x = 270, y = 0, xend = 270, yend = -4.8), color="black", size=1) +
    # Point1
    new_scale_color() +
    geom_point(data = df2, aes(x=dayJ2, y=depth2, fill=type2, color=type2),  size = 4, pch=21, stroke =1.8) +# 40
          scale_fill_manual(name="", values=c( "#3b5a9d", "#FEC287FF"))  +
          scale_color_manual(name="", values = c( "grey79", "black")) +

    
    xlab("Day of year") +
    
  theme_light() +
    theme(legend.position = "none", plot.margin = margin(t=5, b=0, r=20, l=5, "mm")) +  
    theme(axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 9, color = "black"),
        axis.text.y = element_text(size = 9, color = "black"))

)
```

# Legend
```{r}
( legend = Temp2 %>%
    filter(id=="151") %>%
    ggplot(., mapping = aes(x = julianTemp, y = TempC)) + 
    geom_line(color='#006D5B')+
   scale_y_continuous(
      name = (expression("Temperature " ( degree*C))), breaks = c(-10,-5, 0, 10, 20),
      sec.axis = dup_axis( name="", breaks=c(0,-1.6,-3.46, -4.8), labels = c("0","5", "38", "132"))) +
  geom_segment(aes(x = 150, y = 0, xend = 150, yend = -3.46), color="black", size=1) +
  geom_segment(aes(x = 270, y = 0, xend = 270, yend = -4.8), color="black", size=1) +
    # Point1
    new_scale_color() +
    geom_point(data = df2, aes(x=dayJ2, y=depth2, fill=type2, color=type2),  size = 4, pch=21, stroke =1.8) +# 40
          scale_fill_manual(name="", values=c( "#3b5a9d", "#FEC287FF"))  +
          scale_color_manual(name="", values = c( "grey79", "black")) +

    
    xlab("Day of year") +
    
  theme_light() +
    theme(legend.position = "right", 
          plot.margin = margin(t=5, b=0, r=20, l=5, "mm")) 

)
```


# Combine

```{r}

pt_legend = cowplot::get_legend(legend)
```
```{r}
( combineFig = cowplot::plot_grid(plot151Thaw2, plot150Active2, ncol=1, labels = "AUTO") + 
  draw_label("Depth (cm)", color = "black", x=  .87, y=.225, vjust= 1.5,size = 10, angle = 0)  + 
  draw_label("Depth (cm)", color = "black", x=  .86, y=.72, vjust= 1.5,size = 10, angle = 0) +
  draw_grob(pt_legend, x = .4, y=0, width=1, height=1, scale = .8)
  )
```

```{r}
ggsave("../figures/TempMeasuredPredicted.png", plot = combineFig, width = 6, height =6, units = c("in"), dpi=600, bg = "white" )
```