---
title: "FigureMapPlusPhotos"
author: "Anna Talucci"
date: "2023-10-02"
output: html_document
---


# clear environment
```{r}
rm(list=ls())
```


# Overview

Notes:
[Better colors for Mapping](https://www.esri.com/arcgis-blog/products/js-api-arcgis/mapping/better-colors-for-better-mapping/)
# Packages
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
library(scales)
library(ggrepel)
library(patchwork)
```

# Projection
[Some projection info]https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection)
```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"


# Data
## Points
```{r}
photoPts = st_read("../data/photoPts/photoPts.shp", "photoPts")
```

```{r}
( photoPts = photoPts %>% rename(aLabs = label) )
```

```{r}
PtsEcoSumXY = st_read("../data/allPoints/PtsEcoSumXY.shp", "PtsEcoSumXY")
```

```{r}
PtsEcoSumXY
```

```{r}
unique(PtsEcoSumXY$ECO_NAME)
```

## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/shapefiles/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```


## permafrost
```{r}
permafrost = st_read("../data/shapefiles/permafrost/permaice.shp", "permaice")
```

```{r}
head(permafrost)
```

## WWF
```{r}
wwf = st_read("../data/wwfEcozones/wwf_terr_ecos.shp", "wwf_terr_ecos")
```


### WWF Data 
```{r}
wwf 
```

### Select Biomes
```{r}
( eco = 
    wwf %>% 
    filter(REALM %in% c("PA", "NA")) %>% 
    dplyr::select(OBJECTID, ECO_NAME, REALM, BIOME, ECO_NUM, ECO_ID, geometry) %>% 
    filter(BIOME %in% c(6, 11)) %>% 
    mutate(color_eco = ifelse(ECO_NAME == "Arctic coastal tundra" , '#a6cee3',
                        ifelse(ECO_NAME ==  "Arctic foothills tundra" , '#1f78b4',
                        ifelse(ECO_NAME == "Beringia lowland tundra" , '#b2df8a',
                        ifelse(ECO_NAME == "East Siberian taiga"  , '#33a02c',
                        ifelse(ECO_NAME == "Interior Alaska-Yukon lowland taiga" , '#fb9a99',
                        ifelse(ECO_NAME == "Interior Yukon-Alaska alpine tundra" , '#e31a1c',
                        ifelse(ECO_NAME == "Muskwa-Slave Lake forests"  , '#fdbf6f',
                        ifelse(ECO_NAME == "Northeast Siberian taiga" , '#ff7f00',
                        ifelse(ECO_NAME == "Northern Canadian Shield taiga" , '#cab2d6',
                        ifelse(ECO_NAME == "Northwest Territories taiga" , '#6a3d9a',
                        ifelse(ECO_NAME == "Yamal-Gydan tundra" , '#ffff99', '#CCCCCC'))))))))))))
)
```

 [1] "Arctic coastal tundra"               "Arctic foothills tundra"             "Beringia lowland tundra"            
 [4] "East Siberian taiga"                 "Interior Alaska-Yukon lowland taiga" "Interior Yukon-Alaska alpine tundra"
 [7] "Muskwa-Slave Lake forests"           "Northeast Siberian taiga"            "Northern Canadian Shield taiga"     
[10] "Northwest Territories taiga"         "Yamal-Gydan tundra"                 

```{r}
ecoPalette = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99', '#CCCCCC')
```
```{r}
plot(st_geometry(eco))
```


## Arctic Circle 
Arctic Circle shapefile came from [Natural Earth](https://www.naturalearthdata.com/downloads/110m-physical-vectors/110m-geographic-lines/)

```{r}
circle = st_read("../data/ne_110m_geographic_lines/ne_110m_geographic_lines.shp", "ne_110m_geographic_lines") 
```
```{r}
arctic = circle %>% filter(name=="Arctic Circle")
```



# Filter shapefilres

## Countries

```{r}
target = c("Asia", "Europe", "North America")
```

Filter for all
```{r}
nh  = wrld %>% 
  filter(CONTINENT %in% target) %>%
  dplyr::select( OBJECTID:FORMAL_EN, NAME_EN, ISO_A2, CONTINENT:REGION_WB, Shape_Leng:geometry) 
```


```{r}
nh
```

## Permafrost

```{r}
permafrost1 = permafrost %>% 
  filter(EXTENT %in% c("C", "D", "S")) 
```

# Reproject

```{r}
nh_pp = st_transform(nh, crs = polarProj)
artic_pp = st_transform(arctic, crs = polarProj)
eco_pp = st_transform(eco, crs = polarProj)
photo_pp = st_transform(photoPts, crs = polarProj)
```

permafrost1_pp = st_transform(permafrost1, crs = polarProj)
PtsEcoSumXY_pp = st_transform(PtsEcoSumXY, crs = polarProj)

```{r}
nh_pp
eco_pp
```

```{r}
nh_cropped <- st_crop(nh_pp, xmin = -180, xmax = 180,
                                    ymin = 45, ymax = 90)
```

# Palettes

```{r}
ecoPalette = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#CCCCCC')
```


```{r}
permafrost_palette = c('#71A6D1', '#99C4E1', '#C0E2F0')
```

```{r}
brltnd_palette = c('#4a6741', '#3d3021')
```
xlim = c(-3302324, 2312498), ylim = c(-1108650,  3102452) 
ymin: -25426070 xmax: 28628080 ymax: 8108557
 xmin:  ymin:  xmax:  ymax: 
 
 scale_color_manual(values = c("A" ="#fdae61","B" = "#ff6347", "C" = "#fee08b"), labels = c( "A" ="Tweedsmuir fire","B" = "Entiako fire", "C" = "Chelaslie fire" ), name = "Fire perimeters") +

# Maps

## Zoom Level
```{r}
zoom_to <- c(0, 90)  # Geographic North pole

zoom_level <- 3

lon_span <- 360 / 2^zoom_level
lat_span <- 180 / 2^zoom_level
```

```{r}
lon_bounds <- c(zoom_to[1] - lon_span / 2, zoom_to[1] + lon_span / 2)
lat_bounds <- c(zoom_to[2] - lat_span / 2, zoom_to[2] + lat_span / 2)
```
coord_sf(xlim = lon_bounds, ylim = lat_bounds) +
## Display window 
```{r}
# corner points at bottom left and top right
# supply as WGS84 and directly transform to target CRS (Mollweide)
cornerpts <- data.frame(label = c('A', 'B'))
cornerpts$geometry <- st_transform(st_sfc(st_point(c(279.26, 33.92)), st_point(c(102.34, 31.37)),
                                          crs = 4326),
                                   crs = polarProj)
cornerpts <- st_as_sf(cornerpts)
```

```{r}
disp_win_wgs84 <- st_sfc(st_point(c(279.26, 33.92)), st_point(c(102.34, 31.37)), crs = 4326)
disp_win_wgs84

disp_win_trans <- st_transform(disp_win_wgs84, crs = polarProj)
disp_win_trans

disp_win_coord <- st_coordinates(disp_win_trans)
```

```{r}
#### zoom to kamchatka, different projection ####

zoom_to <- c(180, 83)  # ~ center of Kamchatka
zoom_level <- 2

# Lambert azimuthal equal-area projection around center of interest
target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.5)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = target_crs)
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = target_crs)

```

## Overview map by Ecozone
```{r}
( map = ggplot() +
     geom_sf(data = nh_pp, fill = "#F9F6EE", colour="#A9AB9D") +
      geom_sf(data = eco_pp, fill = eco_pp$color_eco, color=eco_pp$color_eco) +
 
 geom_sf(data=artic_pp, fill=NA, colour = "#2F4F4F", lwd=.5, linetype="dashed", alpha=.7) +
new_scale_fill() +
  geom_point(data = photoPts, aes(x=x, y=y,)) +

  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
    #coord_sf(xlim = disp_win_coord[,'X'], ylim = disp_win_coord[,'Y']) +
  #coord_sf(xlim = c(-4002324, 3512498), ylim = c(-3008650,  4002452) ) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#daf5f3")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())
)
```

# Maps with Legend
```{r}
eco_pp
```
```{r}
unique(eco_pp$ECO_NAME)

```

## Select for legend
```{r}
legend_eco = eco_pp %>%filter(ECO_NAME %in% c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Yamal-Gydan tundra", "Yukon Interior dry forests" ))
```

```{r}
( legend = ggplot() +
     geom_sf(data = nh_pp, fill = "#F9F6EE", colour="#A9AB9D") +
   
     geom_sf(data = legend_eco, aes(fill = ECO_NAME, color=ECO_NAME))+
 
   scale_fill_manual(values=ecoPalette, name="Ecozones", labels = c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Yamal-Gydan tundra", "Other")) +
    scale_color_manual(values=ecoPalette, name="Ecozones", labels = c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Yamal-Gydan tundra", "Other")) +


  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
    #coord_sf(xlim = disp_win_coord[,'X'], ylim = disp_win_coord[,'Y']) +
  #coord_sf(xlim = c(-4002324, 3512498), ylim = c(-3008650,  4002452) ) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom",
         legend.text=element_text(size=6),
        legend.margin = margin(0, 0, 0, 0),
        legend.title=element_text(size=7),
         legend.key=element_blank(),
        legend.key.size = unit(5, 'mm'),
        axis.title.x = element_blank(), 
        axis.title.y = element_blank()) +
  guides(color = guide_legend(title.position = "top", ncol=3,bycol=TRUE, override.aes = list(size = 4)),
         fill = guide_legend(title.position = "top", ncol=3,bycol=TRUE, override.aes = list(size = 4)))
)
```

scale_fill_identity(values=eco_pp$color_eco, name="Ecozones", labels = c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Yamal-Gydan tundra", "Other")) +
    scale_color_identity(values=eco_pp$color_eco, name="Ecozones", labels = c("Arctic coastal tundra", "Arctic foothills tundra", "Beringia lowland tundra", "East Siberian taiga", "Interior Alaska-Yukon lowland taiga", "Interior Yukon-Alaska alpine tundra", "Muskwa-Slave Lake forests", "Northeast Siberian taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", "Yamal-Gydan tundra", "Other")) +

# Trying to clip map to circle
```{r eval=FALSE, include=FALSE}
x_lines <- seq(-120,180, by = 60)

( p1 = ggplot() +
  geom_sf(data = nh_pp, fill = "#F9F6EE", colour="#A9AB9D") +
      geom_sf(data = eco_pp, fill = eco_pp$color_eco, color=eco_pp$color_eco) +
 geom_sf(data=artic_pp, fill=NA, colour = "#2F4F4F", lwd=.5, linetype="dashed", alpha=.7) +
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())
 )

p1 + coord_sf(crs = sf::st_crs("ESRI:102016"))
```


# Grid map and photos

## legend
```{r}
legendBot = cowplot::get_legend(legend)
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/legendBotPhotoMap.png", plot = legendBot, dpi = 600,  bg='white')
```
## Lines
```{r}
lineA1 = data.frame(x1 = 1.3, x2 = 3.0, y1 = 4.3, y2 = 4.05)
lineA2 = data.frame(x1 = 1.8, x2 = 3.0, y1 = 4.3, y2 = 4.05)
```

```{r}
lineB1 = data.frame(x1 = 1.2, x2 = 2.1, y1 = 3.5, y2 = 3.5)
lineB2 = data.frame(x1 = 1.2, x2 = 2.15, y1 = 3.1, y2 = 3.4)
```

```{r}
lineC1 = data.frame(x1 = 4.3, x2 = 4.8, y1 = 2.6, y2 = 2.5)
lineC2 = data.frame(x1 = 4.3, x2 = 4.8, y1 = 2.6, y2 = 1.7)
lineC3 = data.frame(x1 = 4.3, x2 = 4.8, y1 = 2.6, y2 = .9)
```

## figure
```{r fig.height=7, fig.width=6}
fig1 = ggdraw(xlim = c(0, 6), ylim = c(0, 6)) +
    draw_plot(map, x = .9, y = 1, width = 4.5, height = 4) + 
  
  #YKD
  
  geom_segment(data =lineA1, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#b2df8a', linewidth= .5) +
  geom_segment(data =lineA2, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#b2df8a', linewidth = .5) +
  draw_image("../images/ForMapFigure/Unburned_YKD_Baillargeon.JPG", scale = 0.8, x = 0, y = 4.2, width = 1.5, height = 1) +
  draw_image("../images/ForMapFigure/15Burn_YKD_Baillargeon.JPG", scale = 0.8, x = 1.3, y = 4.2, width = 1.5, height = 1) +
    draw_label("A1", x = .25, y = 5, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  draw_label("A2", x = 1.6, y = 5, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  
  # Canada
  
  geom_segment(data =lineB1, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#fdbf6f', linewidth= .5) +
  geom_segment(data =lineB2, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#fdbf6f', linewidth = .5) +
    draw_image("../images/ForMapFigure/KakisaB2014Quad1.JPG", scale = 0.8, x = 0, y = 3.3, width = 1.5, height = 1) +
    draw_image("../images/ForMapFigure/CamsellUBQuad2.JPG", scale = 0.8, x = 0, y = 2.4, width = 1.5, height = 1) +
     draw_label("B1", x = .35, y = 4.1, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  draw_label("B2", x = .35, y = 3.1, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  #Siberia
    geom_segment(data =lineC1, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#ffff99', linewidth= .5) +
  geom_segment(data =lineC2, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#ffff99', linewidth = .5) +
  geom_segment(data =lineC3, aes(x = x1, y = y1, xend = x2, yend = y2), colour = '#ffff99', linewidth = .5) +
    draw_image("../images/ForMapFigure/DJI_0999.JPG", scale = 0.8, x = 4.5, y = 1.8, width = 1.5, height = 1) +
    draw_image("../images/ForMapFigure/DJI_0031.JPG", scale = 0.8, x = 4.5, y = 1.1, width = 1.5, height = 1) +
draw_image("../images/ForMapFigure/DJI_0015.JPG", scale = 0.8, x = 4.5, y = .4, width = 1.5, height = 1.) +
  
   draw_label("C1", x = 4.75, y = 2.5, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  draw_label("C2", x = 4.75, y = 1.8, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  draw_label("C3", x = 4.75, y = .9, fontfamily = "sans", fontface = "bold", color = "black", size = 10) +
  #legend
  draw_image("../images/legendBotPhotoMap.PNG", scale = 0.8, x =1.6, y = .15, width = 3, height = 2) 

ggsave("../figures/PhotoMap.png", plot = fig1, width = 6, height = 6, dpi = 600, bg='white')
```



  geom_segment(data =line1, aes(x = x1, y = y1, xend = x2, yend = y2), colour = "#000080", linewidth= 1, alpha = .5) +
  geom_segment(data =line2, aes(x = x1, y = y1, xend = x2, yend = y2), colour = "#000080", linewidth = 1, alpha = .5) 
  draw_label("A", x = 0.3, y = 6.8, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
   draw_label("B", x = 0.3, y = 3.3, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
   draw_label("C", x = 3.0, y = 3.3, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
```{r}
ggsave("../figures/PhotoMap.png", plot = fig1, width = 6, height = 6, dpi = 600, bg='white')
```