---
title: "Manies Clean to Point"
author: "Anna Talucci"
date: '2022-12-11'
output: html_document
---


# clear environment
```{r}
rm(list=ls())
```

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Manies
```{r}
f8 <- read.csv(f[8], header = TRUE)
f8
```
```{r}
str(f8)
```
# fix incorrect logical columns
```{r}
f8$slope <- read.csv(f[8], header = TRUE, colClasses = "character")[,20]

f8$organic_depth <- read.csv(f[8], header = TRUE, na.strings = "unk")[,15]
```

```{r}
f8_dates = f8 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) %>%
  dplyr::select(-notes)

f8_dates
```






```{r}
f8_dates %>% distinct(site_id, thaw_active, julianDate, year, date)
```
Apr 1  to September 30 (91-273)
```{r}
ggplot(f8_dates, aes(x=lat, y=long, color=thaw_active)) +
  geom_jitter()
```

# Split by active & thaw
```{r}
f8_thaw = f8_dates %>% 
  filter(thaw_active == "T") 

f8_thaw

f8_active = f8_dates %>% 
  filter(thaw_active == "A") 

f8_active
```

# Save clean data
```{r}
write_csv(f8_active, "../outputs/CleanActive/ManiesCleanActive.csv")
```

# Create points shapefile

## For F8
sf::st_as_sf(dd, coords = c("x","y"))
```{r}
f8_thaw_pts = st_as_sf(f8_thaw, coords = c("long","lat"), crs = 4326, remove = FALSE)
f8_active_pts = st_as_sf(f8_active, coords = c("long","lat"), crs = 4326, remove = FALSE)
f8_pts = st_as_sf(f8_dates, coords = c("long","lat"), crs = 4326, remove = FALSE)
```


### write to shapefile
```{r}
st_write(f8_thaw_pts, "../outputs/ThawPoints/Manies_thaw_pts.shp", driver="ESRI Shapefile")
st_write(f8_active_pts, "../outputs/ActivePoints/Manies_active_pts.shp", driver="ESRI Shapefile")
st_write(f8_pts, "../outputs/allPoints/Manies_pts.shp", driver="ESRI Shapefile")
```