---
title: "Subset data for ALT dates"
author: "Anna Talucci"
date: "2023-04-27"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview

Standardize thaw depth measures to ALT. 


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
library(gt)
library(gtsummary)
```
# Data

## Standardized measures
```{r}
allDataALT = read_csv("../data/Standardize/AllDataStandardizeALT.csv")
```
"../outputs/Standardize/AllDataStandardizeALT.csv")

## vector of data files
```{r}
f <- list.files(path = "../data/era5Clean/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Individual files
```{r}
( f2001 <- read_csv(f[18]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
)
```
```{r}
f2005 <- read_csv(f[19]) %>% mutate(altDate=minDate-14)%>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2006 <- read_csv(f[20]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2007 <- read_csv(f[21]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2008 <- read_csv(f[22]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2009 <- read_csv(f[23]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2010 <- read_csv(f[24]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2011 <- read_csv(f[25]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2012 <- read_csv(f[26]) %>% mutate(altDate=minDate-14)%>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1)%>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2013 <- read_csv(f[27]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2014 <- read_csv(f[28]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2015 <- read_csv(f[29]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2016 <- read_csv(f[30]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2017 <- read_csv(f[31]) %>% mutate(altDate=minDate-14)%>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2018 <- read_csv(f[32]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
f2019 <- read_csv(f[33]) %>% mutate(altDate=minDate-14) %>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
```

```{r}
(
  f2020 <- read_csv(f[34]) %>% mutate(altDate=minDate-14)%>% 
  dplyr::select(id, plot_id:cntry_c, day:id, minDate:altDate) %>%
  group_by(id) %>%
  slice(1) %>%
  mutate(plot_id = as.character(plot_id))
)
```

# Combine

```{r}
(
  combo = dplyr::bind_rows(f2001, f2005, f2006, f2007, f2008, f2009, f2010, f2011, f2012, f2013, f2014, f2015, f2016, f2017, f2018, f2019, f2020)
)
```

# Calculate difference between dates 45179
```{r}
(
  diff = combo %>% 
    filter(altDate > 0) %>%
  mutate(dateDiff = altDate-julinDt) %>%
    dplyr::select(id, plot_id, site_id, year, minDate, altDate, dateDiff)
)

```

# Combine Data
```{r}
allDataALT
```
# Merge data frames
```{r}
( allData1 = merge(allDataALT, diff, by = c("id", 'plot_id', 'site_id', 'year')) )
```

# Difference in Measure and adjust ALT
```{r}
( allData2 = allData1 %>% 
    mutate(diffALT = AdjALT - thw_dpt))
```

```{r}
write.csv(allData2,file = "../outputs/Standardize/AllDataStandardizeALTwithDates.csv")
```


# Tables
```{r}
( table1 = allData2 %>% 
  dplyr::select(cntry_c, thw_ctv, dateDiff) %>% 
    tbl_strata(
    strata = cntry_c,
    ~.x %>%
  tbl_summary(  
    by=thw_ctv,
    type = all_continuous() ~ "continuous2",       
    statistic = all_continuous() ~ c(
      "{mean} ({sd})",                             
      "{median} ({p25}, {p75})",                   
      "{min}, {max}"),                            
    label  = list(dateDiff   ~ "difference in measured and calculated date"))) %>%
  as_gt() %>%
  gt::tab_header(title = "Difference in measured and calculated date",
                 subtitle = "Summary statistics by Country")
) 
```

```{r eval=FALSE, include=FALSE}
gt::gtsave(table1, filename = "../outputs/Tables/DatesTable1.html")
```


```{r}
( table2 = allData2 %>% 
  dplyr::select(cntry_c, thw_ctv, diffALT) %>% 
    tbl_strata(
    strata = cntry_c,
    ~.x %>%
  tbl_summary(  
    by=thw_ctv,
    type = all_continuous() ~ "continuous2",       
    statistic = all_continuous() ~ c(
      "{mean} ({sd})",                             
      "{median} ({p25}, {p75})",                   
      "{min}, {max}"),                            
    label  = list(diffALT   ~ "difference in measured and calculated ALT"))) %>%
  as_gt() %>%
  gt::tab_header(title = "Difference in measured and calculated ALT",
                 subtitle = "Summary statistics by Country and whether it was original measured as Active or Thaw")
) 
```

```{r eval=FALSE, include=FALSE}
gt::gtsave(table2, filename = "../outputs/Tables/AdjustedMeasureDiffTable2.html")
```
allData %>% 
  dplyr::select(brn_nbr, cntry_c, AdjALT) %>% 
  tbl_strata(
    strata = cntry_c,
    ~.x %>%
      tbl_summary(
        by = brn_nbr,
    type = all_continuous() ~ "continuous2",       
    statistic = all_continuous() ~ c(
      "{mean} ({sd})",                             
      "{median} ({p25}, {p75})",                   
      "{min}, {max}"), 
    label = list(AdjALT   ~ "Standardized Active Layer Thickness"))) %>%
  as_gt() %>%
  gt::tab_header(title = "Standardized measures of Active Layer Thickness",
                 subtitle = "Summary statistics")
```{r}
( table2 = diff %>% 
  dplyr::select(thw_ctv, diff) %>% 
    tbl_strata(
    strata = cntry_c,
    ~.x %>%
  tbl_summary(  
    by=thw_ctv,
    type = diff ~ "continuous2",       
    statistic = diff ~ c(
      "{mean} ({sd})",                             
      "{median} ({p25}, {p75})",                   
      "{min}, {max}"),                            
    label  = list(diff   ~ "difference in measured and calculated date"))) %>%
  as_gt() %>%
  gt::tab_header(title = "Difference in measured and calculated date",
                 subtitle = "Summary statistics by Active or Thaw measure")
) 
```

```{r eval=FALSE, include=FALSE}
gt::gtsave(table2, filename = "../outputs/Tables/DatesThawActiveTable2.html")
```

```{r}
( meandiff = combo %>% 
    filter(altDate > 0) %>%
  mutate(diff = altDate-julinDt) %>%
    group_by(last_nm, site_id, plot_id, brn_nbr, cntry_c) %>%
  summarize(meanDiff = mean(diff))
)
```

```{r}
( ca = diff  %>% filter(cntry_c =="CA")
  
)
ru = diff  %>% filter(cntry_c =="RU")
us = diff  %>% filter(cntry_c =="US")
```

```{r}
# Color based on value
color <- ifelse(ca$meanDiff < 0, "#f1a340", "#998ec3")

ggplot(ca, aes(x = reorder(id, meanDiff), y = meanDiff)) +
  geom_bar(stat = "identity",
           show.legend = FALSE,
           fill = color,     
           color = "white") +
  geom_hline(yintercept = 0, color = 1, lwd = 0.2) +
  #geom_text(aes(label = group, # Text with groups
                #hjust = ifelse(value < 0, 1.5, -1),
               # vjust = 0.5), size = 2.5) +
  xlab("Group") +
  ylab("difference") +
  #scale_y_continuous(breaks = seq(-2, 2, by = 1),limits = c(-2.5, 2.5)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_blank(),  # Remove Y-axis texts
        axis.ticks.y = element_blank(), # Remove Y-axis ticks
        panel.grid.major.y = element_blank()) + # Remove horizontal grid
  facet_wrap(~cntry_c, scales = "free_y")
```