---
title: " Gibson/Olefeldt Clean to Point"
author: "Anna Talucci"
date: '2022-12-11'
output: html_document
---

Gibson/Olefeldt


# clear environment
```{r}
rm(list=ls())
```

# Overview

Adapt script form M. Loranty data_cleaning, to only clean and not aggregate data. 

Use data from csv files in Google Drive

What this script does:


# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```
# Projection

WGS 84 need for gee
```{r}
wgs_proj =  "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "
```

# Data

# vector of data files
```{r}
f <- list.files(path = "../data/GoogleDriveCsvFolder/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```



## Olefeldt
```{r}
f11 <- read.csv(f[11], header = TRUE)
f11
```

# fix incorrect logical columns
```{r}
# fix incorrect logical columns
f11$slope <- read.csv(f[11], header = TRUE, colClasses = "character")[,20]

# remove non-numeric characters from numeric vars
# note we're loosing info on where Organic Layer Thickness is in excess of the entered value
f11$organic_depth <- as.numeric(gsub(">", "",f11$organic_depth))

f11 %>% distinct(site_id,year,month,day,fire_id,burn_unburn)

```

```{r}
str(f11)
```

```{r}
f11_dates = f11 %>%
  mutate(date = as.Date(with(., paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  mutate(julianDate = yday(date)) %>%
  dplyr::select(-fire_id)

f11_dates
```


```{r}
f11_dates %>% filter_all(any_vars(is.na(.))) 
```



```{r}
f11_dates %>% distinct(julianDate, year, date)
```

# Split by active & thaw
```{r}
f11_thaw = f11_dates %>% 
  filter(thaw_active == "T") 

f11_thaw

f11_active = f11_dates %>% 
  filter(thaw_active == "A") 

f11_active
```
```{r}
unique(f11_active$date)
```

```{r}
unique(f11_thaw$date)
```
# Save clean data
```{r}
write_csv(f11_active, "../outputs/CleanActive/OlefeldtCleanActive.csv")
```

# Create points shapefile

## For F11
sf::st_as_sf(dd, coords = c("x","y"))
```{r}
f11_thaw_pts = st_as_sf(f11_thaw, coords = c("long","lat"), crs = 4326, remove = FALSE)
f11_active_pts = st_as_sf(f11_active, coords = c("long","lat"), crs = 4326, remove = FALSE)
```


### write to shapefile
```{r}
st_write(f11_thaw_pts, "../outputs/ThawPoints/Olefeldt_thaw_pts.shp", driver="ESRI Shapefile")
st_write(f11_active_pts, "../outputs/ActivePoints/Olefeldt_active_pts.shp", driver="ESRI Shapefile")
```