---
title: "Soil and Air Temperature"
author: "Anna Talucci"
date: "2023-03-15"
output: html_document
---

# Overview

Notes form email with Mike:
Regarding the date - I think the approach you describe using climate data probably makes sense, though the maximum may be reached before mean temps drop below freezing. This paper indicates that max thaw occurs during the last two weeks in August in Utqiagvik, and I think average temps are still above freezing at this point. In my limited experience in Cherskiy, thaw depth tends to plateau around early Sept when temps are still above freezing as well. It might be the case that max thaw occurs roughly when air and soil temps are approximately similar, and once air temps get lower there can be bottom-up freezing as the soil loses heat to the atmosphere. In Cherskiy soil temps never typically do not exceed 4C. Sorry, but of a stream of consciousness here. 

PCN Burned/Unburned Paired contributed data Lat lon
Baillargeon     US        61.2704 -163.236
Buma            US        65.71578 -149.226
 Rocha      US 68.99 -150.28 

Douglas      US 64.71577 -148.0096
Frost      US 61.2711 -162.6291
Gaglioti      US 67.99161 -162.0033 
 Manies      US 63.8474 -145.717
 
 O'Donnell      US 65.57626 -148.9595 
 Natali      US 68.9908 -150.27 
 
 Dieleman      CA 62.732 -115.651 
 Gibson      CA 62.24611 -122.6356
 
 Paulson      RU 68.10885 161.3355 

 Sizov    RU 65.85096 74.38171 
 Delcourt-Veraverbeke      RU 62.07247 125.8988 
# Packages
```{r}
library(tidyverse)
library(lubridate)
```
# Graph theme
```{r}

reg_theme = theme_bw() + theme(legend.position = "bottom") +
  theme(plot.margin = unit(c(t = 0.3, r = 0.3, b = 0.3, l = 0.1), "cm")) +
  theme(axis.title.x = element_text(size = 11, hjust = 0.5, vjust = -0.1),
        axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

```

# Data 

```{r}
soil = read.csv("../data/soilAirTemp/soil_temp.csv")
air = read.csv("../data/soilAirTemp/air_temp.csv")
site = read.csv("../data/soilAirTemp/siteinfo.csv")
```


```{r}
head(soil)
head(air)
head(site)
```

```{r}
unique(soil$site_id)
unique(air$site_id)
```
```{r}
unique(soil$st_depth)
```

# Join

```{r}
( air1 = site %>% full_join(y=air, by="site_id") )
( soil1 = site %>% full_join(y=soil, by="site_id") )
```

# Filter by hemisphere

```{r}
( air_west = air1 %>% filter(lon < 0) %>% drop_na(air_t) %>% mutate(combo = as.integer(interaction(lat,lon,drop = TRUE))))
( soil_west = soil1 %>% filter(lon < 0) %>% drop_na(soil_t) %>% mutate(combo = as.integer(interaction(lat,lon,drop = TRUE))))
( air_east = air1 %>% filter(lon > 0) %>% drop_na(air_t) %>% mutate(combo = as.integer(interaction(lat,lon,drop = TRUE))))
( soil_east = soil1 %>% filter(lon > 0) %>% drop_na(soil_t)%>% mutate(combo = as.integer(interaction(lat,lon,drop = TRUE))))
```

# Unique pairs of lat lon
## East hemisphere
 Paulson      RU 68.10885 161.3355 (combo 11)
 Sizov    RU 65.85096 74.38171 
 Delcourt-Veraverbeke      RU 62.07247 125.8988 
```{r}
unique(air_east[c( "site_id","lat","lon", "combo")])
```

## Western Hemisphere
Canda
 Dieleman      CA 62.732 -115.651 (combo 110)
 Gibson      CA 62.24611 -122.6356 (combo 104)
 
US
Baillargeon     US        61.2704     -163.236
Buma            US        65.71578    -149.226        (combo 60)
Rocha           US        68.99       -150.28         (combo 45)
Douglas         US        64.71577    -148.0096       (combo 85)
Frost           US        61.2711     -162.6291
Gaglioti        US        67.99161    -162.0033 
Manies          US        63.8474     -145.717
O'Donnell       US        65.57626    -148.9595 
Natali          US        68.9908     -150.27         (combo 46)
```{r}
unique(air_west[c( "site_id","lat","lon", "combo")])
```
# Filter based on matching temperature for date
## Eastern
```{r}
soil_east %>% 
    filter(combo ==11) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(st_depth, doy_st) %>%
  filter(soil_t==soil_t)


air_east %>% 
  filter(combo ==11) %>%
  filter(doy_ai %in% (200:285)) %>%
  filter(air_t<0)

```



# Summarize & Plot
```{r}
theme_set(theme_minimal())
```
## Eastern Hemisphere


```{r}
( soilEastLat1 = soil_east %>% 
    filter(combo ==11) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airEastLat1 = air_east %>% 
  filter(combo ==11) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilEastLat1, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airEastLat1, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=240, colour = "#fdae61") + 
  geom_vline(xintercept=254, colour="#fdae61") + 
  geom_vline(xintercept=264, colour="#fdae61") + 
  geom_hline(yintercept=0, colour="#d7191c") + 
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") + 
  geom_text(aes(x=240, y= 20, label="Russia (68.75556, 161.45570)"), colour = "black") +
  theme(legend.position = "bottom")


```


## Western Hemisphere
### Canada
111, 110
```{r}
target_ca = c(104)
```

```{r}
( soilWestLat1 = soil_west %>% 
    filter(combo %in% target_ca) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airWestLat1 = air_west %>% 
  filter(combo %in% target_ca) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilWestLat1, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airWestLat1, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=215, colour = "#fdae61") + 
  geom_vline(xintercept=225, colour = "#fdae61") + 
  geom_vline(xintercept=257, colour = "#fdae61") + 
  geom_vline(xintercept=264, colour="#fdae61") +
  geom_hline(yintercept=0, colour="#d7191c") + 
  facet_wrap(~combo, nrow = 3) +
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=240, y= 20, label=""), colour = "black") +
  theme(legend.position = "bottom") +
  geom_text(aes(x=240, y= 20, label="Canada (62.24500, -121.57100)"), colour = "black") 
	


```

### Canada
```{r}
target_us = c(45, 46, 60, 85)
```

```{r}
( soilWestLat2 = soil_west %>% 
    filter(combo ==45) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airWestLat2 = air_west %>% 
  filter(combo ==45) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilWestLat2, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airWestLat2, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=215, colour = "#fdae61") + # Aug 3
  geom_vline(xintercept=260, colour="#fdae61") + #Sept 24
  geom_hline(yintercept=0, colour="#d7191c") + 
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=215, y=-5,label="Aug 3"), colour = "black", angle=90, vjust = -0.5, size=2) +
  geom_text(aes(x=260, y=-5, label="Sept 17"),colour = "black", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=240, y= 20, label=""), colour = "black") +
  theme(legend.position = "right")


```

```{r}
( soilWestLat3 = soil_west %>% 
    filter(combo ==46) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airWestLat3 = air_west %>% 
  filter(combo ==46) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilWestLat3, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airWestLat3, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=215, colour = "#fdae61") + # Aug 3
  geom_vline(xintercept=260, colour="#fdae61") + #Sept 24
  geom_hline(yintercept=0, colour="#d7191c") + 
  facet_wrap(~combo, nrow = 2) +
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=215, y=-5,label="Aug 3"), colour = "black", angle=90, vjust = -0.5, size=2) +
  geom_text(aes(x=260, y=-5, label="Sept 17"),colour = "black", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=240, y= 20, label=""), colour = "black") +
  theme(legend.position = "right")


```

```{r}
( soilWestLat4 = soil_west %>% 
    filter(combo ==60) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airWestLat4 = air_west %>% 
  filter(combo ==60) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilWestLat4, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airWestLat4, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=215, colour = "#fdae61") + # Aug 3
  geom_vline(xintercept=260, colour="#fdae61") + #Sept 24
  geom_hline(yintercept=0, colour="#d7191c") + 
  facet_wrap(~combo, nrow = 2) +
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=215, y=-5,label="Aug 3"), colour = "black", angle=90, vjust = -0.5, size=2) +
  geom_text(aes(x=260, y=-5, label="Sept 17"),colour = "black", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=240, y= 20, label=""), colour = "black") +
  theme(legend.position = "right")


```

```{r}
( soilWestLat5 = soil_west %>% 
    filter(combo ==85) %>%
    filter(st_depth < 60) %>%
  filter(doy_st %in% (200:285)) %>%
  group_by(combo, st_depth, doy_st) %>%
  summarize(mean_soil_t = mean(soil_t))
)

( airWestLat5 = air_west %>% 
  filter(combo ==85) %>%
  filter(doy_ai %in% (200:285)) %>%
  group_by(combo, doy_ai) %>%
  summarize(mean_air_t = mean(air_t))
)
```

```{r}

  ggplot(data = soilWestLat5, aes(x=doy_st, y=mean_soil_t, color=as.factor(st_depth)))  +
  geom_line() +
  geom_line(data = airWestLat5, aes(x=doy_ai, y=mean_air_t),color = "black") +
  geom_vline(xintercept=215, colour = "#fdae61") + # Aug 3
  geom_vline(xintercept=260, colour="#fdae61") + #Sept 24
  geom_hline(yintercept=0, colour="#d7191c") + 
  facet_wrap(~combo, nrow = 2) +
  scale_color_viridis_d(name = "Depth (cm)") +
  xlab("Julian date") +
  ylab("Temperature (Celecius)") +
  geom_text(aes(x=215, y=-5,label="Aug 3"), colour = "black", angle=90, vjust = -0.5, size=2) +
  geom_text(aes(x=260, y=-5, label="Sept 17"),colour = "black", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=240, y= 20, label=""), colour = "black") +
  theme(legend.position = "right")


```

